---
title: "Fitted Models"
author: "Loïc Pages"
date: "`r Sys.Date()`"
output:
  pdf_document: default
word_document: default
html_document:
  df_print: paged
---
  
```{r,include=FALSE}
rm(list=ls())
library(knitr)
library(spaMM)
library(tidyverse)
library(splines)
library(foreach)
library(doParallel)
library(patchwork)

setwd("/media/loic/Commun/0Travail/Stage 2025 ISEM/Code")

IPM_data <- read.csv("newdata.csv")

centauree_data <- IPM_data[!is.na(IPM_data$Size0Mars) & !is.na(IPM_data$Age),]
centauree_data$Age[centauree_data$Age > 8] <- 8

spaMM.options(separation_max=70)
```

```{r,include=FALSE}
annees <- 1995:2022
populations <- c("E2","E1","Au","Po","Pe","Cr")

taille_range <- seq(0.5, 25, by = 0.5) 
age_range <- 1:8

fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = taille_range,
  Age = age_range
)

fake_data <- fake_data %>% 
  mutate(Individu = row_number()) 

fake_data1 <- fake_data[fake_data$Age==1,]
fake_data2 <- fake_data[fake_data$Age>1,]
fake_data3 <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = taille_range,
  Age = age_range,
  Size1 = taille_range)
```

# Seedlings survival

```{r,include=FALSE}
survdata <- centauree_data[centauree_data$Flowering!=1,]
survdata1 <- survdata[survdata$Age==1,]
```

Models selected by AIC
```{r}
ASurvglm11 <- fitme(Survie ~ 1+ bs(Size0Mars,df=4,degree=2) + (Size0Mars|year),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

ASurvglm12 <- fitme(Survie ~ 1+ bs(Size0Mars,df=4,degree=2) + (Size0Mars|year)+ (1|Pop),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

ASurvglm13 <- fitme(Survie ~ 1+ bs(Size0Mars,df=4,degree=2) + (1|year) + (1|Pop),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

ASurvglm14 <- fitme(Survie ~ 1+ bs(Size0Mars,df=4,degree=2) + (1|year),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

ASurvglm15 <- fitme(Survie ~ 1+ bs(Size0Mars,df=5,degree=3) + (Size0Mars|year),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")
```

Models selected by BIC
```{r}
BSurvglm11 <- fitme(Survie ~ 1+ Size0Mars + (1|year) + (1|Pop),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

BSurvglm12 <- fitme(Survie ~ 1+ Size0Mars + (1|year),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

BSurvglm13 <- fitme(Survie ~ 1 + Size0Mars + (Size0Mars|year) + (1|Pop),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

BSurvglm14 <- fitme(Survie ~ 1+ Size0Mars + (Size0Mars|year),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")

BSurvglm15 <- fitme(Survie ~ 1+ Size0Mars + (1|year) + (Size0Mars|Pop),
                  family=binomial,
                  data=survdata1,
                  method="PQL/L")
```

```{r,include=FALSE}
ASurvpredict1 <- predict(ASurvglm11, newdata = fake_data1)[,1]
ASurvpredict2 <- predict(ASurvglm12, newdata = fake_data1)[,1]
ASurvpredict3 <- predict(ASurvglm13, newdata = fake_data1)[,1]
ASurvpredict4 <- predict(ASurvglm14, newdata = fake_data1)[,1]
ASurvpredict5 <- predict(ASurvglm15, newdata = fake_data1)[,1]

BSurvpredict1 <- predict(BSurvglm11, newdata = fake_data1)[,1]
BSurvpredict2 <- predict(BSurvglm12, newdata = fake_data1)[,1]
BSurvpredict3 <- predict(BSurvglm13, newdata = fake_data1)[,1]
BSurvpredict4 <- predict(BSurvglm14, newdata = fake_data1)[,1]
BSurvpredict5 <- predict(BSurvglm15, newdata = fake_data1)[,1]
```

```{r,include=FALSE}
plot_survie1 <- function(data = fake_data1, prediction, var, c1, valc1, fact) {
  data %>%
    mutate(surv_predi = prediction) %>%
    filter(!!sym(c1) == valc1) %>%
    ggplot(aes(x = .data[[var]], y = surv_predi)) +
    geom_count(data=survdata1,aes(y=Survie),alpha = 0.5,show.legend = FALSE) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    ylim(0, 1)+
    xlim(0,15)
}

plot_survie1bis <- function(data = fake_data1, prediction, var, c1, valc1, fact) {
  data %>%
    mutate(surv_predi = prediction) %>%
    filter(!!sym(c1) == valc1) %>%
    ggplot(aes(x = .data[[var]], y = surv_predi)) +
    geom_count(data=survdata1,aes(y=Survie),alpha = 0.5,show.legend = FALSE) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    ylim(0, 1)+
    xlim(0,15)+
   scale_color_viridis_d(option = "plasma")

}
```


## As a function of size

### See year effect (population fixed)
```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "Pop"
valc1 <- "Au"
fact <- "year"
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie1(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1,  fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie1(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1,  fact=fact)
)
```

### See population effect (year fixed)
```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "year"
valc1 <- 2000
fact <- "Pop"
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie1bis(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1,  fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie1bis(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, fact=fact),
  plot_survie1bis(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1,  fact=fact)
)
```


# Rosette survival

```{r,include=FALSE}
survdata2 <- survdata[survdata$Age>1,]
```

AIC
```{r}
ASurvglm21 <- fitme(Survie ~ 1+ bs(Size0Mars,df=3,degree=2) + 
                      bs(Age,degree=3,knots=6.5) + (Age|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

ASurvglm22 <- fitme(Survie ~ 1+  bs(Size0Mars,df=3,degree=2) + 
                      bs(Age,degree=3,knots=6.5) + (1|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

ASurvglm23 <- fitme(Survie ~ 1+ bs(Size0Mars,df=4,degree=2) + 
                      bs(Age,degree=3,knots=6.5) + (Age|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

ASurvglm24 <- fitme(Survie ~ 1 + poly(Size0Mars,4) + bs(Age,degree=3,knots=6.5) 
                    + (Age|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

ASurvglm25 <- fitme(Survie ~ 1+ bs(Size0Mars,df=4,degree=2) + 
                      bs(Age,degree=3,knots=6.5) + (1|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")
```

BIC
```{r}
BSurvglm21 <- fitme(Survie ~ 1+ bs(Size0Mars,df=3,degree=2) + (Age|year) + 
                      (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

BSurvglm22 <- fitme(Survie ~ 1+ poly(Size0Mars,3) + (Age|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

BSurvglm23 <- fitme(Survie ~ 1 + bs(Size0Mars,df=3,degree=2) + (Age|year) + 
                      (Size0Mars|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

BSurvglm24 <- fitme(Survie ~ 1+ bs(Size0Mars,df=3,degree=2) + (Age|year) + 
                      (Age|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")

BSurvglm25 <- fitme(Survie ~ 1+ poly(Size0Mars,2) + (Age|year) + (1|Pop),
                  family=binomial,
                  data=survdata2,
                  method="PQL/L")
```


```{r,include=FALSE}
ASurvpredict1 <- predict(ASurvglm21, newdata = fake_data2)[,1]
ASurvpredict2 <- predict(ASurvglm22, newdata = fake_data2)[,1]
ASurvpredict3 <- predict(ASurvglm23, newdata = fake_data2)[,1]
ASurvpredict4 <- predict(ASurvglm24, newdata = fake_data2)[,1]
ASurvpredict5 <- predict(ASurvglm25, newdata = fake_data2)[,1]

BSurvpredict1 <- predict(BSurvglm21, newdata = fake_data2)[,1]
BSurvpredict2 <- predict(BSurvglm22, newdata = fake_data2)[,1]
BSurvpredict3 <- predict(BSurvglm23, newdata = fake_data2)[,1]
BSurvpredict4 <- predict(BSurvglm24, newdata = fake_data2)[,1]
BSurvpredict5 <- predict(BSurvglm25, newdata = fake_data2)[,1]
```


```{r,include=FALSE}
plot_survie <- function(data = fake_data2, prediction, var, c1, valc1 = 1, c2, valc2 = "Au", fact) {
  data %>%
    mutate(surv_predi = prediction) %>%
    filter(!!sym(c1) == valc1, !!sym(c2) == valc2) %>%
    ggplot(aes(x = .data[[var]], y = surv_predi)) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend=FALSE) +
    theme_minimal() +
    ylim(0, 1)
}

plot_surviebis <- function(data = fake_data2, prediction, var, c1, valc1 = 1, c2, valc2 = "Au", fact) {
  data %>%
    mutate(surv_predi = prediction) %>%
    filter(!!sym(c1) == valc1, !!sym(c2) == valc2) %>%
    ggplot(aes(x = .data[[var]], y = surv_predi)) +
    geom_count(data=survdata2,aes(x=Age,y=Survie),alpha = 0.5) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend=FALSE) +
    theme_minimal() +
    ylim(0, 1)
}

plot_survie2 <- function(data = fake_data2, prediction, var, c1, valc1 = 1, c2, valc2 = "Au", fact) {
  data %>%
    mutate(surv_predi = prediction) %>%
    filter(!!sym(c1) == valc1, !!sym(c2) == valc2) %>%
    ggplot(aes(x = .data[[var]], y = surv_predi)) +
    geom_count(data=survdata2[survdata2$Age %in% c(2,8),],aes(y=Survie,color = as.factor(.data[[fact]])),alpha = 0.3,show.legend = FALSE) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    ylim(0, 1)+
      scale_color_brewer(palette = "Spectral", direction = -1)
}
```

## As a function of size

### See effect of age (population and year fixed)
```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "year"
c2 <- "Pop"
valc2 <- "Au"
fact <- "Age"
valc1 <- 2000
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie2(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie2(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie2(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

### See year effect (population fixed and age=5)
```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "Age"
c2 <- "Pop"
valc2 <- "Au"
fact <- "year"
valc1 <- 5
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

### See population effect (year fixed and age=5)
```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "Age"
c2 <- "year"
valc2 <- 2000
fact <- "Pop"
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

BIC 
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_survie(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_survie(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

## As a function of age (size=5)

### See year effect (population fixed)

```{r,include=FALSE}
var <- "Age"
c1 <- "Size0Mars"
c2 <- "Pop"
valc2 <- "Au"
fact <- "year"
valc1 <- 5
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_surviebis(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_surviebis(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact),
  plot_surviebis(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact)
)
```

<!-- ### See population effect (year fixed) -->

<!-- ```{r,include=FALSE} -->
<!-- var <- "Age" -->
<!-- c1 <- "Size0Mars" -->
<!-- c2 <- "year" -->
<!-- valc2 <- 2000 -->
<!-- fact <- "Pop" -->
<!-- ``` -->

<!-- AIC -->
<!-- ```{r,warning=FALSE,echo=FALSE} -->
<!-- wrap_plots( -->
<!--   plot_surviebis(prediction = ASurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = ASurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = ASurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = ASurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = ASurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact) -->
<!-- ) -->
<!-- ``` -->

<!-- BIC -->
<!-- ```{r,warning=FALSE,echo=FALSE} -->
<!-- wrap_plots( -->
<!--   plot_surviebis(prediction = BSurvpredict1, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = BSurvpredict2, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = BSurvpredict3, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = BSurvpredict4, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact), -->
<!--   plot_surviebis(prediction = BSurvpredict5, var=var, c1=c1, valc1=valc1, c2=c2, valc2=valc2, fact=fact) -->
<!-- ) -->
<!-- ``` -->



# Flowering probability


AIC
```{r,include=FALSE}
AFlowglm1 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,2) + (Age|Pop),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

AFlowglm2 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,4) + (Age|Pop),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

AFlowglm3 <- fitme(Flowering ~ 1+ poly(Size0Mars,4) + poly(Age,2) + (Age|Pop),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

AFlowglm4 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,2) + (Age|Pop)+ (1|year),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

AFlowglm5 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,3) + (Age|Pop),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")
```

BIC
```{r,include=FALSE}
BFlowglm1 <- fitme(Flowering ~ 1 + poly(Size0Mars,3) + poly(Age,2) + (Age|Pop),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

BFlowglm2 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,2) + (Age|Pop) + 
                     (1|year),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

BFlowglm3 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,2) + (Age|Pop) + (Size0Mars|year),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

BFlowglm4 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,2) + (Age|Pop) + (Age|year),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")

BFlowglm5 <- fitme(Flowering ~ 1+ poly(Size0Mars,3) + poly(Age,2) + 
                     (Size0Mars + Age|Pop),
                  family=binomial,
                  data=centauree_data,
                  method="PQL/L")
```


```{r,include=FALSE}
AFlowpredict1 <- predict(AFlowglm1, newdata = fake_data)[,1]
AFlowpredict2 <- predict(AFlowglm2, newdata = fake_data)[,1]
AFlowpredict3 <- predict(AFlowglm3, newdata = fake_data)[,1]
AFlowpredict4 <- predict(AFlowglm4, newdata = fake_data)[,1]
AFlowpredict5 <- predict(AFlowglm5, newdata = fake_data)[,1]

BFlowpredict1 <- predict(BFlowglm1, newdata = fake_data)[,1]
BFlowpredict2 <- predict(BFlowglm2, newdata = fake_data)[,1]
BFlowpredict3 <- predict(BFlowglm3, newdata = fake_data)[,1]
BFlowpredict4 <- predict(BFlowglm4, newdata = fake_data)[,1]
BFlowpredict5 <- predict(BFlowglm5, newdata = fake_data)[,1]
```


```{r,include=FALSE}
plot_Flowering <- function(data = fake_data, prediction) {
  data %>%
  mutate(flow_predi = prediction) %>%
  group_by(Size0Mars, Age) %>%
  summarise(flow_predi = mean(flow_predi),
            .groups = "drop") %>%
  ggplot(aes(x = Size0Mars, y = flow_predi)) +
  geom_count(data=centauree_data,aes(x=Size0Mars,y=Flowering,col=as.factor(Age)),alpha = 0.6,show.legend = FALSE) +
  geom_line(aes(color = as.factor(Age)),size=0.75,show.legend = FALSE) +
  theme_bw()+
  ylim(0, 1) +
  labs(x = "Size",
      y = "Flowering probability",
      color = "Age")+
  scale_color_brewer(palette = "Spectral", direction = -1)
}

plot_Floweringbis <- function(data = fake_data, prediction) {
  data %>%
  mutate(flow_predi = prediction) %>%
  group_by(Size0Mars, Age) %>%
  summarise(flow_predi = mean(flow_predi),
            .groups = "drop") %>%
  ggplot(aes(x = Size0Mars, y = flow_predi)) +
  geom_count(data=centauree_data,aes(x=Size0Mars,y=Flowering,col=as.factor(Age)),alpha = 0.6,show.legend = FALSE) +
  geom_line(aes(color = as.factor(Age)),size=0.75) +
  theme_bw()+
  ylim(0, 1) +
  labs(x = "Size",
      y = "Flowering probability",
      color = "Age")+
  scale_color_brewer(palette = "Spectral", direction = -1)
}

plot_Flowering2 <- function(data = fake_data, prediction, var, fact) {
  data %>%
    mutate(flow_predi = prediction) %>%
    filter(Size0Mars==10) %>% 
    group_by(!!sym(var), !!sym(fact)) %>%
     summarise(flow_predi = mean(flow_predi),
            .groups = "drop") %>%
    ggplot(aes(x = .data[[var]], y = flow_predi)) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    scale_color_viridis_d(option = "plasma")+
    ylim(0, 1)
}

plot_Flowering2bis <- function(data = fake_data, prediction, var, fact) {
  data %>%
    mutate(flow_predi = prediction) %>%
    filter(Size0Mars==10) %>% 
    group_by(!!sym(var), !!sym(fact)) %>%
    summarise(flow_predi = mean(flow_predi),
            .groups = "drop") %>%
    ggplot(aes(x = .data[[var]], y = flow_predi)) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    ylim(0, 1)
}

```

## As a function of size

### See effect of age

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_Flowering(prediction = AFlowpredict1),
  plot_Flowering(prediction = AFlowpredict2),
  plot_Flowering(prediction = AFlowpredict3),
  plot_Flowering(prediction = AFlowpredict4),
  plot_Floweringbis(prediction = AFlowpredict5)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_Flowering(prediction = BFlowpredict1),
  plot_Flowering(prediction = BFlowpredict2),
  plot_Flowering(prediction = BFlowpredict3),
  plot_Flowering(prediction = BFlowpredict4),
  plot_Floweringbis(prediction = BFlowpredict5)
)
```


## As a function of age (size fixed)

### See population effect (mean over the years)

```{r,include=FALSE}
var <- "Age"
fact <- "Pop"
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_Flowering2(prediction = AFlowpredict1, var=var, fact=fact),
  plot_Flowering2(prediction = AFlowpredict2, var=var, fact=fact),
  plot_Flowering2(prediction = AFlowpredict3, var=var, fact=fact),
  plot_Flowering2(prediction = AFlowpredict4, var=var, fact=fact),
  plot_Flowering2(prediction = AFlowpredict5, var=var, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_Flowering2(prediction = BFlowpredict1, var=var, fact=fact),
  plot_Flowering2(prediction = BFlowpredict2, var=var, fact=fact),
  plot_Flowering2(prediction = BFlowpredict3, var=var, fact=fact),
  plot_Flowering2(prediction = BFlowpredict4, var=var, fact=fact),
  plot_Flowering2(prediction = BFlowpredict5, var=var, fact=fact)
)
```

### See year effect (mean over the populations)
 

```{r,include=FALSE}
var <- "Age"
fact <- "year"
```

AIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_Flowering2bis(prediction = AFlowpredict1, var=var, fact=fact),
  plot_Flowering2bis(prediction = AFlowpredict2, var=var, fact=fact),
  plot_Flowering2bis(prediction = AFlowpredict3, var=var, fact=fact),
  plot_Flowering2bis(prediction = AFlowpredict4, var=var, fact=fact),
  plot_Flowering2bis(prediction = AFlowpredict5, var=var, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_Flowering2bis(prediction = BFlowpredict1, var=var, fact=fact),
  plot_Flowering2bis(prediction = BFlowpredict2, var=var, fact=fact),
  plot_Flowering2bis(prediction = BFlowpredict3, var=var, fact=fact),
  plot_Flowering2bis(prediction = BFlowpredict4, var=var, fact=fact),
  plot_Flowering2bis(prediction = BFlowpredict5, var=var, fact=fact)
)
```

# Growth

```{r,include=FALSE}
growthdata <- centauree_data[!is.na(centauree_data$Size1Mars), ]
growthdata <- growthdata[growthdata$Size1Mars != 0, ]
```


AIC
```{r}
AGrowthglm1 <- fitme(Size1Mars ~ 1 + poly(Size0Mars,3) + 
                       bs(Age,degree=2,knots=6.5) + (Size0Mars+Age|year) + 
                       (1|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

AGrowthglm2 <- fitme(Size1Mars ~ 1 + poly(Size0Mars,3) + 
                       bs(Age,degree=2,knots=6.5) + (Size0Mars+Age|year) +
                       (Size0Mars|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

AGrowthglm3 <- fitme(Size1Mars ~ 1 + poly(Size0Mars,3) + 
                       bs(Age,degree=2,knots=6.5) + (Size0Mars+Age|year) +
                       (Size0Mars+Age|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

AGrowthglm4 <- fitme(Size1Mars ~ 1 + bs(Size0Mars,df=5,degree=3) + 
                       bs(Age,degree=2,knots=6.5) + (Size0Mars+Age|year) + 
                       (1|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

AGrowthglm5 <- fitme(Size1Mars ~ 1 + bs(Size0Mars,df=5,degree=3) +
                       bs(Age,degree=2,knots=6.5) + (Size0Mars+Age|year) +
                       (Size0Mars|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)
```


BIC
```{r}
BGrowthglm1 <- fitme(Size1Mars ~ 1 + poly(Size0Mars,3) + poly(Age,2) +
                       (Size0Mars+Age|year) + (1|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

BGrowthglm2 <- fitme(Size1Mars ~ 1 + poly(Size0Mars,3) + poly(Age,2) +
                       (Size0Mars+Age|year) + (Size0Mars+Age|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

BGrowthglm3 <- fitme(Size1Mars ~ 1 + bs(Size0Mars,df=3,degree=2) + poly(Age,2) +
                       (Size0Mars+Age|year) + (1|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

BGrowthglm4 <- fitme(Size1Mars ~ 1 + bs(Size0Mars,df=3,degree=2) + poly(Age,2) +
                       (Size0Mars+Age|year) + (Size0Mars|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)

BGrowthglm5 <- fitme(Size1Mars ~ 1 + bs(Size0Mars,df=3,degree=2) + poly(Age,2) +
                       (Size0Mars+Age|year) + (Size0Mars+Age|Pop),
                    resid.model = ~ log(Size0Mars)+log(Age),
                    data=growthdata)
```

```{r,include=FALSE}
AGrowthpredict1 <- predict(AGrowthglm1, newdata = fake_data)[,1]
AGrowthpredict2 <- predict(AGrowthglm2, newdata = fake_data)[,1]
AGrowthpredict3 <- predict(AGrowthglm3, newdata = fake_data)[,1]
AGrowthpredict4 <- predict(AGrowthglm4, newdata = fake_data)[,1]
AGrowthpredict5 <- predict(AGrowthglm5, newdata = fake_data)[,1]

BGrowthpredict1 <- predict(BGrowthglm1, newdata = fake_data)[,1]
BGrowthpredict2 <- predict(BGrowthglm2, newdata = fake_data)[,1]
BGrowthpredict3 <- predict(BGrowthglm3, newdata = fake_data)[,1]
BGrowthpredict4 <- predict(BGrowthglm4, newdata = fake_data)[,1]
BGrowthpredict5 <- predict(BGrowthglm5, newdata = fake_data)[,1]
```


```{r}
plot_growth1 <- function(data = fake_data, prediction, var, c1, c2, valc1=1, fact) {
  data %>%
    mutate(size1predi = prediction) %>%
    group_by(!!sym(var),!!sym(fact)) %>% 
    filter(!!sym(c1) == valc1) %>%
    summarise(size1predi = mean(size1predi),
            .groups = "drop") %>%
    ggplot(aes(x = .data[[var]], y = size1predi)) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    geom_abline()+
    theme_minimal()
}

plot_growth2 <- function(data = fake_data, prediction, var, c1, c2, valc1=1, fact) {
  data %>%
    mutate(size1predi = prediction) %>%
    group_by(!!sym(var),!!sym(fact)) %>%
    filter(!!sym(c1) == valc1) %>%
    summarise(size1predi = mean(size1predi),
            .groups = "drop") %>%
    ggplot(aes(x = .data[[var]], y = size1predi)) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    geom_abline()+
    theme_minimal()+
    scale_color_viridis_d(option = "plasma")
}

plot_growth3 <- function(data = fake_data, prediction, var, c1, c2, valc1=1, fact) {
  data %>% mutate(sdgrow = prediction) %>% 
  group_by(Size0Mars, Age) %>%
  summarise(sd_predi = mean(sdgrow),
            .groups = "drop") %>%
  ggplot(aes(x = Size0Mars, y = sd_predi)) +
  geom_line(aes(color = as.factor(Age))) +
  theme_bw()+
  labs(x = "Size(t)",
      y = "Growth residual variance",
      fill = "Age",
      color = "Age")+
    scale_color_brewer(palette = "Spectral", direction = -1)
}
```

## Size at t+1 as a function of size at t (age 1, 4, 8)

### See year effect (population fixed)

```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "Age"
c2 <- "Pop"
fact <- "year"
```

AIC
```{r,warning=FALSE,echo=FALSE}
valc1 <- 1
wrap_plots(
  plot_growth1(prediction = AGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)

valc1 <- 4
wrap_plots(
  plot_growth1(prediction = AGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)


valc1 <- 8
wrap_plots(
  plot_growth1(prediction = AGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = AGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
valc1 <- 1
wrap_plots(
  plot_growth1(prediction = BGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)

valc1 <- 4
wrap_plots(
  plot_growth1(prediction = BGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)


valc1 <- 8
wrap_plots(
  plot_growth1(prediction = BGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth1(prediction = BGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)
```

#### See population effect (year fixed)

```{r,include=FALSE}
var <- "Size0Mars"
c1 <- "Age"
c2 <- "year"
fact <- "Pop"
```

AIC
```{r,warning=FALSE,echo=FALSE}
valc1 <- 1
wrap_plots(
  plot_growth2(prediction = AGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)

valc1 <- 4
wrap_plots(
  plot_growth2(prediction = AGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)


valc1 <- 8
wrap_plots(
  plot_growth2(prediction = AGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = AGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE}
valc1 <- 1
wrap_plots(
  plot_growth2(prediction = BGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)

valc1 <- 4
wrap_plots(
  plot_growth2(prediction = BGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)


valc1 <- 8
wrap_plots(
  plot_growth2(prediction = BGrowthpredict1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth2(prediction = BGrowthpredict5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)
```

## Residual variance

```{r,include=FALSE}
Arespred1 <- get_residVar(AGrowthglm1,newdata=fake_data)
Arespred2 <- get_residVar(AGrowthglm2,newdata=fake_data)
Arespred3 <- get_residVar(AGrowthglm3,newdata=fake_data)
Arespred4 <- get_residVar(AGrowthglm4,newdata=fake_data)
Arespred5 <- get_residVar(AGrowthglm5,newdata=fake_data)

Brespred1 <- get_residVar(BGrowthglm1,newdata=fake_data)
Brespred2 <- get_residVar(BGrowthglm2,newdata=fake_data)
Brespred3 <- get_residVar(BGrowthglm3,newdata=fake_data)
Brespred4 <- get_residVar(BGrowthglm4,newdata=fake_data)
Brespred5 <- get_residVar(BGrowthglm5,newdata=fake_data)
```

AIC
```{r,echo=FALSE}
wrap_plots(
  plot_growth3(prediction = Arespred1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Arespred2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Arespred3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Arespred4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Arespred5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)
```

BIC
```{r,echo=FALSE}
wrap_plots(
  plot_growth3(prediction = Brespred1, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Brespred2, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Brespred3, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Brespred4, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact),
  plot_growth3(prediction = Brespred5, var=var, c1=c1, valc1=valc1, c2=c2, fact=fact)
)
```



# Number of capitula (fecundity)

```{r,include=FALSE}
cptldata <- centauree_data[centauree_data$Flowering!=0,]
```

AIC
```{r}
ACptlglm1 <- fitme(log(Capitule) ~ 1 + Size0Mars + (Age|year), 
                  data=cptldata)

ACptlglm2 <- fitme(log(Capitule) ~ 1 + Size0Mars + (Age|year) + (1|Pop), 
                  data=cptldata)

ACptlglm3 <- fitme(log(Capitule) ~ 1 + poly(Size0Mars,2) + (Age|year), 
                  data=cptldata)

ACptlglm4 <- fitme(log(Capitule) ~ 1 + Size0Mars + Age +(Age|year), 
                  data=cptldata)

ACptlglm5 <- fitme(log(Capitule) ~ 1 + Size0Mars + (1|year), 
                  data=cptldata)
```

BIC
```{r}
BCptlglm1 <- fitme(log(Capitule) ~ 1 + Size0Mars + (Age|year), 
                  data=cptldata)

BCptlglm2 <- fitme(log(Capitule) ~ 1 +Size0Mars + (Age|year) + (1|Pop), 
                  data=cptldata)

BCptlglm3 <- fitme(log(Capitule) ~ 1 + Size0Mars + (1|year), 
                  data=cptldata)

BCptlglm4 <- fitme(log(Capitule) ~ 1 + Size0Mars, 
                  data=cptldata)

BCptlglm5 <- fitme(log(Capitule) ~ 1 + Size0Mars + (1|year) + (1|Pop), 
                  data=cptldata)
```


```{r,include=FALSE}
ACptlpredict1 <- predict(ACptlglm1, newdata = fake_data)[,1]
ACptlpredict2 <- predict(ACptlglm2, newdata = fake_data)[,1]
ACptlpredict3 <- predict(ACptlglm3, newdata = fake_data)[,1]
ACptlpredict4 <- predict(ACptlglm4, newdata = fake_data)[,1]
ACptlpredict5 <- predict(ACptlglm5, newdata = fake_data)[,1]

BCptlpredict1 <- predict(BCptlglm1, newdata = fake_data)[,1]
BCptlpredict2 <- predict(BCptlglm2, newdata = fake_data)[,1]
BCptlpredict3 <- predict(BCptlglm3, newdata = fake_data)[,1]
BCptlpredict4 <- predict(BCptlglm4, newdata = fake_data)[,1]
BCptlpredict5 <- predict(BCptlglm5, newdata = fake_data)[,1]
```


```{r,include=FALSE}
plot_capitule <- function(data = fake_data, prediction, var, fact) {
  data %>%
    mutate(cptl_predi = exp(prediction)) %>%
    group_by(!!sym(var),!!sym(fact)) %>% 
    summarise(cptl_predi = mean(cptl_predi)) %>%
    ggplot(aes(x = .data[[var]], y = cptl_predi)) +
    geom_point(data = cptldata, aes(y = Capitule), alpha=0.6)+
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    ylim(0,100)
}

plot_capitule2 <- function(data = fake_data, prediction, var, fact) {
  data %>%
    mutate(cptl_predi = exp(prediction)) %>%
    group_by(!!sym(var),!!sym(fact)) %>% 
    summarise(cptl_predi = mean(cptl_predi)) %>%
    ggplot(aes(x = .data[[var]], y = cptl_predi)) +
    geom_point(data = cptldata, aes(y = Capitule), alpha=0.6)+
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    theme_minimal() +
    scale_color_viridis_d(option = "plasma")+
    ylim(0,100)
}
```


## As a function of size

### See year effect

```{r,include=FALSE}
var <- "Size0Mars"
fact <- "year"
```

AIC
```{r,warning=FALSE,echo=FALSE,message=FALSE}
wrap_plots(
  plot_capitule(prediction = ACptlpredict1, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict2, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict3, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict4, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict5, var=var, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE,message=FALSE}
wrap_plots(
  plot_capitule(prediction = BCptlpredict1, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict2, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict3, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict4, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict5, var=var, fact=fact)
)
```

### See population effect

```{r,include=FALSE}
var <- "Size0Mars"
fact <- "Pop"
```

AIC
```{r,warning=FALSE,echo=FALSE,message=FALSE}
wrap_plots(
  plot_capitule2(prediction = ACptlpredict1, var=var, fact=fact),
  plot_capitule2(prediction = ACptlpredict2, var=var, fact=fact),
  plot_capitule2(prediction = ACptlpredict3, var=var, fact=fact),
  plot_capitule2(prediction = ACptlpredict4, var=var, fact=fact),
  plot_capitule2(prediction = ACptlpredict5, var=var, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE,message=FALSE}
wrap_plots(
  plot_capitule2(prediction = BCptlpredict1, var=var, fact=fact),
  plot_capitule2(prediction = BCptlpredict2, var=var, fact=fact),
  plot_capitule2(prediction = BCptlpredict3, var=var, fact=fact),
  plot_capitule2(prediction = BCptlpredict4, var=var, fact=fact),
  plot_capitule2(prediction = BCptlpredict5, var=var, fact=fact)
)
```

## As a function of age

### See year effect

```{r,include=FALSE}
var <- "Age"
fact <- "year"
```

AIC
```{r,warning=FALSE,echo=FALSE,message=FALSE}
wrap_plots(
  plot_capitule(prediction = ACptlpredict1, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict2, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict3, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict4, var=var, fact=fact),
  plot_capitule(prediction = ACptlpredict5, var=var, fact=fact)
)
```

BIC
```{r,warning=FALSE,echo=FALSE,message=FALSE}
wrap_plots(
  plot_capitule(prediction = BCptlpredict1, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict2, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict3, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict4, var=var, fact=fact),
  plot_capitule(prediction = BCptlpredict5, var=var, fact=fact)
)
```



# Seedling size

```{r,include=FALSE}
plantule_data <- centauree_data[centauree_data$Age==1,]
```


Same with AIC and BIC
```{r,include=FALSE}
Pltglm1 <- fitme(Size0Mars ~ 1 + (1|year) + (1|Pop) + (1|Pop:year), 
                 data=plantule_data,
                 family = Gamma(log))

Pltglm2 <- fitme(Size0Mars ~ 1 + (1|Pop) + (1|Pop:year), 
                 data=plantule_data,
                 family = Gamma(log))

Pltglm3 <- fitme(Size0Mars ~ 1 + (1|year) + (1|Pop:year), 
                 data=plantule_data,
                 family = Gamma(log))

Pltglm4 <- fitme(Size0Mars ~ 1 + (1|Pop:year), 
                 data=plantule_data,
                 family = Gamma(log))

Pltglm5 <- fitme(Size0Mars ~ 1 + (1|year) + (1|Pop), 
                 data=plantule_data,
                 family = Gamma(log))
```


```{r,include=FALSE}
Pltpredict1 <- predict(Pltglm1, newdata = fake_data)[,1]
Pltpredict2 <- predict(Pltglm2, newdata = fake_data)[,1]
Pltpredict3 <- predict(Pltglm3, newdata = fake_data)[,1]
Pltpredict4 <- predict(Pltglm4, newdata = fake_data)[,1]
Pltpredict5 <- predict(Pltglm5, newdata = fake_data)[,1]
```


```{r,include=FALSE}
plot_plantule <- function(data = fake_data, prediction, var, fact) {
  data %>%
    mutate(plt_predi = prediction) %>%
    ggplot(aes(x = .data[[var]], y = plt_predi)) +
    geom_line(aes(color = as.factor(.data[[fact]])),show.legend = FALSE) +
    labs(y="Taille des plantules")+
    scale_color_viridis_d(option = "plasma")+
    theme_minimal()
}

plot_plantule1 <- function(data = fake_data, prediction, var, fact) {
  data %>%
    mutate(plt_predi = prediction) %>%
    ggplot(aes(x = .data[[var]], y = plt_predi)) +
    geom_line(aes(color = as.factor(.data[[fact]]))) +
    labs(y="Taille des plantules")+
    scale_color_viridis_d(option = "plasma")+
    theme_minimal()
}

plot_plantule2 <- function(data = fake_data, prediction) {
  data %>%
    mutate(plt_predi = prediction) %>%
    ggplot(aes(x = plt_predi)) +
    stat_bin(binwidth = 0.25,fill="grey",color="black")+
    labs(x="Taille des plantules")+
    theme_minimal()
}
```


## As a function of year

```{r,include=FALSE}
var <- "year"
fact <- "Pop"
```

```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_plantule(prediction = Pltpredict1, var=var, fact=fact),
  plot_plantule(prediction = Pltpredict2, var=var, fact=fact),
  plot_plantule(prediction = Pltpredict3, var=var, fact=fact),
  plot_plantule(prediction = Pltpredict4, var=var, fact=fact),
  plot_plantule1(prediction = Pltpredict5, var=var, fact=fact)
)
```


## Density

```{r,warning=FALSE,echo=FALSE}
wrap_plots(
  plot_plantule2(prediction = Pltpredict1),
  plot_plantule2(prediction = Pltpredict2),
  plot_plantule2(prediction = Pltpredict3),
  plot_plantule2(prediction = Pltpredict4),
  plot_plantule2(prediction = Pltpredict5)
)
```