---
title: "Models Selection"
author: "Loïc Pages"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

# Introduction

```{r}
rm(list=ls())
library(knitr)
library(spaMM)
library(tidyverse)
library(splines)
library(foreach)
library(doParallel)


setwd("/media/loic/Commun/0Travail/Stage 2025 ISEM/Code")

IPM_data <- read.csv("newdata.csv")


centauree_data <- IPM_data[!is.na(IPM_data$Size0Mars) & !is.na(IPM_data$Age),]
centauree_data$Age[centauree_data$Age > 8] <- 8

spaMM.options(separation_max=70)
```



# Models

Modèle IPM
$$N(y,t+1)=\int (F_a(x,y,t)+P_a(x,y,t))N(x,t)dx$$
avec $x$ la taille à $t$, $y$ la taille à $t+1$, $a$ l'age et $N$ la taille de la population

La probabilité qu'un individu d'age a et de taille x au temps t devienne un individu d'age a+1 et de taille y à t+1 est :
$$P_a(x,y,t)=s_a(x)(1-f_a(x))g_a(x,y)$$
et la densité d'individus de taille y à t et d'age 1 (plantules) issus d'un individu de taille x et d'age a à t est :
$$F_a(x,y,t)=f_a(x)c_a(x)\omega(y)Estb$$


# Model selection

Méthode : 
- Determine the distribution form (gaussian, binomial, etc)
- Create list of fixed effects
- Create list of random effects
- Crate all sub-model combinations with these effects
- Test AIC (or BIC) and select the best model


Vital rates to model :
Survival probability
Flowering probability
Growth
Fecundity (number of capitula)
Seedling size distribution
Establishment rate


## Create list of fixed and random effects
```{r}
# Random effects
r_effects <- c("year", "Pop")

# Fixed effects on which random effects have an effect
f_effects <- c("1", "Size0Mars", "Age", "Size0Mars + Age")


# Combinations (fixed effects | random effect) 
ran_combinations <- sapply(f_effects, function(feffect) {
  sapply(r_effects, function(reffect) {
    paste0("(", feffect, "|", reffect, ")")
  })
})

# Add individual effect
random_effects <- c(as.vector(ran_combinations),"1","(1|Individu)")
```


```{r}
# Size fixed effect
size_effects <- c("Size0Mars","poly(Size0Mars,2)","bs(Size0Mars,df=4,degree=2)","bs(Size0Mars,df=3,degree=2)",
                   "poly(Size0Mars,3)","bs(Size0Mars,df=5,degree=3)","poly(Size0Mars,4)")

# Age fixed effect
age_effects <- c("Age", "poly(Age,2)", "bs(Age,degree=2,knots=6.5)", "poly(Age,3)", "bs(Age,degree=3,knots=6.5)", "bs(Age,degree=3,knots=c(1.5,6.5))", "poly(Age,4)")

age_effects2 <- c("Age", "poly(Age,2)", "bs(Age,degree=2,knots=6.5)", "poly(Age,3)", "bs(Age,degree=3,knots=6.5)", "poly(Age,4)") #splines with only one knot

age_effects3 <- c("Age", "poly(Age,2)", "poly(Age,3)", "poly(Age,4)") #no splines
```



```{r}
#Fonctions that remove unecessary repetitions 
# Example: if there is 2 random effects (x|year) we remove the combination from the list.
is_valid_combination <- function(comb) {
  comb <- sort(comb)  # Trier pour vérifier l'ordre
  if (sum(grepl("\\|year", comb)) >= 2) return(FALSE)
  if (sum(grepl("\\|Pop", comb)) >= 2) return(FALSE)
  if (sum(grepl("\\|Individu", comb)) >= 2) return(FALSE)
  
  if (!"1" %in% comb) return(FALSE) # 1 for intercept is needed for all combinations
  return(TRUE)
}

# Remove individual effect
individual_combination <- function(comb){
    if ("(1|Individu)" %in% comb) return(FALSE)
  return(TRUE)
}

# Function that removes non-performing combinations where there is a random individual effect (for survival age>1).
is_valid_Individu <- function(comb){
  if ("(1|Individu)" %in% comb && any(grepl("^bs", comb))) return(FALSE)
  if ("(1|Individu)" %in% comb && "(Size0Mars + Age|year)" %in% comb) return(FALSE)
  if ("(1|Individu)" %in% comb && "(Age|year)" %in% comb) return(FALSE)
  if ("(1|Individu)" %in% comb && "(Size0Mars|year)" %in% comb) return(FALSE) 
  if ("(1|Individu)" %in% comb && "(Size0Mars + Age|Pop)" %in% comb) return(FALSE)
  if ("(1|Individu)" %in% comb && "(Age|Pop)" %in% comb) return(FALSE)
  if ("(1|Individu)" %in% comb && "(Size0Mars|Pop)" %in% comb) return(FALSE)
  return(TRUE)
}

# Function that removes combinations where there is a spline function and the random effect individual
is_valid_Individu2 <- function(comb){
  if ("(1|Individu)" %in% comb && any(grepl("^bs", comb))) return(FALSE)
  return(TRUE)
}

# Remove combination where there is age in random effect (for seedling survival)
rage_combination <- function(comb){
  if ("(Age|Pop)" %in% comb) return(FALSE)
  if ("(Size0Mars + Age|Pop)" %in% comb) return(FALSE)
  if ("(Age|year)" %in% comb) return(FALSE)
  if ("(Size0Mars + Age|year)" %in% comb) return(FALSE)
  return(TRUE)
}
```


Generate combinations list
```{r} 
combinations <- list()

# Combinations of all random effects
for (n in 1:length(random_effects)) {
  combinations <- c(combinations, combn(random_effects, n, simplify = FALSE))
}
combinations <- Filter(is_valid_combination, combinations) 


combi <- combinations
# Add size combinations
for (s in size_effects) {
  for (combo in combi) {
    combinations <- c(combinations, list(c(s, combo)))
  }
}

combi_surv1 <- combinations 
combi <- combinations
combi2 <- combinations
combi3 <- combinations

# Add all possible age combinations
for (a in age_effects) {
    for (combo in combinations) {
        combi <- c(combi, list(c(a, combo)))
      }
  }

# Add spline simplified age combinations
for (a in age_effects2) {
    for (combo in combinations) {
        combi2 <- c(combi2, list(c(a, combo)))
      }
}

# Add age combinations without spline
for (a in age_effects3) {
    for (combo in combinations) {
        combi3 <- c(combi3, list(c(a, combo)))
      }
}


validIndividu2 <- Filter(is_valid_Individu,combi2) # Combinaisons simplifiée de l'effet individus

combi_flow <- Filter(individual_combination,combi3) # Flowering combinations

combi_notInd <- Filter(individual_combination,combi) # Combinations without individual effect 

combi_surv1 <- Filter(individual_combination,combi_surv1) 
combi_surv1 <- Filter(rage_combination,combi_surv1) # Seedling survival combinations

combi_surv2 <- Filter(is_valid_Individu,combi) # Rosette survival combinations
```



Autres combinaisons
```{r}
combi_plantule <- c("1","1 + (1|year) + (1|Pop) + (1|Pop:year)","1 + (1|year) + (1|Pop)","1 + (1|year) + (1|Pop:year)","1 + (1|Pop) + (1|Pop:year)","1 + (1|year)","1 + (1|Pop)","1 + (1|Pop:year)")

# Remplace size par log(size)
combi_growth_log <- lapply(combi_notInd, function(x) gsub("Size0Mars", "log(Size0Mars)", x))
combi_growthIndividu_log <- lapply(validIndividu2, function(x) gsub("Size0Mars", "log(Size0Mars)", x))

combi_resvar <- c("1","log(Age)","log(Size0Mars)","log(Size0Mars) + log(Age)")

combi_estb <- list()
for (combo in combi_plantule) {
        combi_estb <- c(combi_estb, list(c("offset(log(Capitule))", combo)))
      }
```

BIC
```{r}
# N the number of subjects
# ntot the total number of observations
extractBIC <- function(fit, ntot, N=1){
  extractAIC(fit)[[2]] +(log(ntot)-2)*DoF(fit)[[3]] + log(N)*DoF(fit)[[1]]
}
```


## Survival probability

```{r}
survdata <- centauree_data[centauree_data$Flowering!=1 & !is.na(centauree_data$Survie),]

survdata1 <- survdata[survdata$Age==1,]
survdata2 <- survdata[survdata$Age>1,]
```

(I recommend to test the more complex model first to see if it is working well)
```{r}
survfit <- fitme(Survie ~ 1 + bs(Size0Mars,df=5,degree=3) + bs(Age,degree=3,knots=6.5)+ (Age+Size0Mars|year) + (Age+Size0Mars|Pop) + (1|Individu),
                     data=survdata2)
summary(survfit)
```


```{r}
sdata <- survdata1
# sdata <- survdata2
ntot <- length(sdata$Individu)
# N <- length(unique(survdata2$Individu))

resp <- "Survie"
combi <- combi_surv1
# combi <- combi_surv2

ls = length(combi)

cl <- makeCluster(10)
registerDoParallel(cl)

result <- foreach(i=1:ls, .packages=c("splines","spaMM")) %dopar% {
  set_effects <- combi[i]
  
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste(resp, "~", effect_formula))

  fit <- fitme(formula, data = sdata, family=binomial, method = "PQL/L")
  
  c(extractAIC(fit)[[2]],extractBIC(fit,ntot),set_effects)
}

stopCluster(cl)

combiresult <- list()
AICresult <- list()
BICresult <- list()
for(i in 1:length(result)){ 
  AICresult[i] <- result[[i]][1]
  BICresult[i] <- result[[i]][2]
  combiresult[i] <- result[[i]][3]
}

posAIC <- order(unlist(AICresult))[1:5]
posBIC <- order(unlist(BICresult))[1:5]

survAIC <- AICresult[posAIC]
survBIC <- BICresult[posBIC]
survmodAIC <- combiresult[posAIC]
survmodBIC <- combiresult[posBIC]

survAIC
survmodAIC

survBIC
survmodBIC 
```


## Floraison

```{r}
i <- 1L
combi_fleur <- list()
# combi_fleur2 <- list()

for (set_effects in combi_flow) {
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste("Flowering", "~", effect_formula))
  if (is_separated.formula(formula, data = centauree_data, method = "PQL/L") == FALSE){
    combi_fleur[[i]] <- set_effects
    i <- i + 1L
  }
}
```

```{r}
fitme(Flowering ~ 1+ poly(Size0Mars,4) + (Size0Mars+Age|year)+poly(Age,3)+(Size0Mars+Age|Pop),
                    family=binomial,
                    data=centauree_data,
                    method="PQL/L")
```


```{r}
resp <- "Flowering"
ntot <- length(centauree_data$Individu)

combi <- combi_fleur
ls <- length(combi)

cl <- makeCluster(10)
registerDoParallel(cl)

result <- foreach(i=1:ls, .packages=c("splines","spaMM")) %dopar% {
  set_effects <- combi[i]
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste(resp, "~", effect_formula))

  fit <- fitme(formula, data = centauree_data, family=binomial, method = "PQL/L")

  c(extractAIC(fit)[[2]],extractBIC(fit,ntot),set_effects)
}

stopCluster(cl)

combiresult <- list()
AICresult <- list()
BICresult <- list()
for(i in 1:length(result)){ 
  AICresult[i] <- result[[i]][1]
  BICresult[i] <- result[[i]][2]
  combiresult[i] <- result[[i]][3]
}


posAIC <- order(unlist(AICresult))[1:5]
posBIC <- order(unlist(BICresult))[1:5]

flowAIC <- AICresult[posAIC]
flowBIC <- BICresult[posBIC]
flowmodAIC <- combiresult[posAIC]
flowmodBIC <- combiresult[posBIC]

flowAIC
flowmodAIC

flowBIC
flowmodBIC 
```

## Croissance
```{r}
growthdata <- centauree_data[!is.na(centauree_data$Size1Mars), ]
growthdata <- growthdata[growthdata$Size1Mars != 0, ]
growthdata <- growthdata[!is.na(growthdata$Age),]
```

### log - log
```{r}
resp <- "log(Size1Mars)"

combi <- combi_growth_log

n <- length(growthdata$Individu)

ls <- length(combi)
lv <- length(combi_resvar)

cl <- makeCluster(10)
registerDoParallel(cl)

result <- foreach(i = 1:ls, .packages=c("splines","spaMM"), .combine = cbind) %dopar% {
    set_effects <- combi[i]

    effect_formula <- paste(unlist(set_effects), collapse = "+")
    formula <- as.formula(paste(resp, "~", effect_formula))
    formula_resvar1 <- as.formula(paste("~", combi_resvar[1]))
    formula_resvar2 <- as.formula(paste("~", combi_resvar[2]))
    formula_resvar3 <- as.formula(paste("~", combi_resvar[3]))
    formula_resvar4 <- as.formula(paste("~", combi_resvar[4]))
    
    fit1 <- fitme(formula, data = growthdata, resid.model = formula_resvar1)
    fit2 <- fitme(formula, data = growthdata, resid.model = formula_resvar2)
    fit3 <- fitme(formula, data = growthdata, resid.model = formula_resvar3)
    fit4 <- fitme(formula, data = growthdata, resid.model = formula_resvar4)

  c(extractAIC(fit1)[[2]],extractAIC(fit2)[[2]],extractAIC(fit3)[[2]],extractAIC(fit4)[[2]])
}

stopCluster(cl)

result2 <- data.frame()
AICresult <- list()
combiresult <- list()

for(i in 1:length(result)){
  for(j in 1:length(combi_resvar)){
    AICresult[j] <- result[[i]][j]
  }
  posAIC <- order(unlist(AICresult))[1]
  combiresult <- result[[i]][length(combi_resvar)+1]
  result2[i,1] <- round(as.numeric(AICresult[posAIC]))
  result2[i,2] <- combiresult
  result2[i,3] <- combi_resvar[posAIC]
}
names(result2)<-c("AIC","Model","ResidVar")

combiresult <- list()
varresult <- list()
AICresult <- list()
for(i in 1:length(result)){
  AICresult[i] <- result2$AIC[i]
  combiresult[i] <- result2$Model[i]
  varresult[i] <- result2$ResidVar[i]
}

posAIC <- order(unlist(AICresult))[1:5]

growthAIC2 <- AICresult[posAIC]
growthmodAIC2 <- combiresult[posAIC]
growthvarAIC2 <- varresult[posAIC]
growthAIC2
growthmodAIC2
growthvarAIC2
```

### taille1 - taille0

```{r}
# Growthglm1 <- fitme(Size1Mars ~ 1 + bs(Size0Mars,df=5,degree=3) + bs(Age,degree=3,knots=6.5)+(Age+Size0Mars|year)+(Age+Size0Mars|Pop),
#                     resid.model =~log(Size0Mars)+log(Age),
#                     data=growthdata)
# summary(Growthglm1)
```

```{r}
resp <- "Size1Mars"

combi <- combi_notInd

n <- length(growthdata$Individu)

ls <- length(combi)
lv <- length(combi_resvar)

cl <- makeCluster(10)
registerDoParallel(cl)

result <- foreach(i = 1:ls, .packages=c("splines","spaMM")) %dopar% {
    set_effects <- combi[i]

    effect_formula <- paste(unlist(set_effects), collapse = "+")
    formula <- as.formula(paste(resp, "~", effect_formula))
    formula_resvar1 <- as.formula(paste("~", combi_resvar[1]))
    formula_resvar2 <- as.formula(paste("~", combi_resvar[2]))
    formula_resvar3 <- as.formula(paste("~", combi_resvar[3]))
    formula_resvar4 <- as.formula(paste("~", combi_resvar[4]))
    
    fit1 <- fitme(formula, data = growthdata, resid.model = formula_resvar1)
    fit2 <- fitme(formula, data = growthdata, resid.model = formula_resvar2)
    fit3 <- fitme(formula, data = growthdata, resid.model = formula_resvar3)
    fit4 <- fitme(formula, data = growthdata, resid.model = formula_resvar4)

  c(extractAIC(fit1)[[2]],extractAIC(fit2)[[2]],extractAIC(fit3)[[2]],extractAIC(fit4)[[2]],effect_formula)
}

stopCluster(cl)

result2 <- data.frame()
AICresult <- list()
combiresult <- list()

for(i in 1:length(result)){
  for(j in 1:length(combi_resvar)){
    AICresult[j] <- result[[i]][j]
  }
  posAIC <- order(unlist(AICresult))[1]
  combiresult <- result[[i]][length(combi_resvar)+1]
  result2[i,1] <- round(as.numeric(AICresult[posAIC]))
  result2[i,2] <- combiresult
  result2[i,3] <- combi_resvar[posAIC]
}
names(result2)<-c("AIC","Model","ResidVar")

combiresult <- list()
varresult <- list()
AICresult <- list()
for(i in 1:length(result)){
  AICresult[i] <- result2$AIC[i]
  combiresult[i] <- result2$Model[i]
  varresult[i] <- result2$ResidVar[i]
}

posAIC <- order(unlist(AICresult))[1:5]

growthAIC2 <- AICresult[posAIC]
growthmodAIC2 <- combiresult[posAIC]
growthvarAIC2 <- varresult[posAIC]
growthAIC2
growthmodAIC2
growthvarAIC2
```


```{r}
resp <- "Size1Mars"

combi <- combi_notInd

n <- length(growthdata$Individu)

ls <- length(combi)
lv <- length(combi_resvar)

cl <- makeCluster(10)
registerDoParallel(cl)

result <- foreach(i = 1:ls, .packages=c("splines","spaMM")) %dopar% {
    set_effects <- combi[i]

    effect_formula <- paste(unlist(set_effects), collapse = "+")
    formula <- as.formula(paste(resp, "~", effect_formula))
    formula_resvar1 <- as.formula(paste("~", combi_resvar[1]))
    formula_resvar2 <- as.formula(paste("~", combi_resvar[2]))
    formula_resvar3 <- as.formula(paste("~", combi_resvar[3]))
    formula_resvar4 <- as.formula(paste("~", combi_resvar[4]))
    
    fit1 <- fitme(formula, data = growthdata, resid.model = formula_resvar1)
    fit2 <- fitme(formula, data = growthdata, resid.model = formula_resvar2)
    fit3 <- fitme(formula, data = growthdata, resid.model = formula_resvar3)
    fit4 <- fitme(formula, data = growthdata, resid.model = formula_resvar4)

  c(extractBIC(fit1,n),extractBIC(fit2,n),extractBIC(fit3,n),extractBIC(fit4,n),effect_formula)
}

stopCluster(cl)

result2 <- data.frame()
BICresult <- list()
combiresult <- list()

for(i in 1:length(result)){
  for(j in 1:length(combi_resvar)){
    BICresult[j] <- result[[i]][j]
  }
  posBIC <- order(unlist(BICresult))[1]
  combiresult <- result[[i]][length(combi_resvar)+1]
  result2[i,1] <- round(as.numeric(BICresult[posBIC]))
  result2[i,2] <- combiresult
  result2[i,3] <- combi_resvar[posBIC]
}
names(result2)<-c("BIC","Model","ResidVar")

combiresult <- list()
varresult <- list()
BICresult <- list()
for(i in 1:length(result)){
  BICresult[i] <- result2$BIC[i]
  combiresult[i] <- result2$Model[i]
  varresult[i] <- result2$ResidVar[i]
}

posBIC <- order(unlist(BICresult))[1:5]

growthBIC2 <- BICresult[posBIC]
growthmodBIC2 <- combiresult[posBIC]
growthvarBIC2 <- varresult[posBIC]
growthBIC2
growthmodBIC2
growthvarBIC2
```
## Nombre de capitules
```{r,message=FALSE}
cptl_data <- centauree_data[centauree_data$Flowering!=0,]

ntot <- length(cptl_data$Individu)
resp <- "log(Capitule)"
combi <- combi_notInd
ls <- length(combi)

cl <- makeCluster(10)
registerDoParallel(cl)

result <- foreach(i = 1:ls,.packages=c("splines","spaMM")) %dopar% {
  set_effects <- combi[i]
  
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste(resp, "~", effect_formula))
  
  fit <- fitme(formula,
               data = cptl_data)
               
  c(extractAIC(fit)[[2]],extractBIC(fit,ntot),set_effects)
}

stopCluster(cl)

combiresult <- list()
AICresult <- list()
BICresult <- list()
for(i in 1:length(result)){ 
  AICresult[i] <- result[[i]][1]
  BICresult[i] <- result[[i]][2]
  combiresult[i] <- result[[i]][3]
}


posAIC <- order(unlist(AICresult))[1:5]
posBIC <- order(unlist(BICresult))[1:5]

cptlAIC <- AICresult[posAIC]
cptlBIC <- BICresult[posBIC]
cptlmodAIC <- combiresult[posAIC]
cptlmodBIC <- combiresult[posBIC]

cptlAIC
cptlmodAIC

cptlBIC
cptlmodBIC 
```

## Taille des plantules

```{r}
resp <- "Size0Mars"
plantule_data <- centauree_data[centauree_data$Age==1,]
ntot <- length(plantule_data$Individu)

ls <- length(combi_plantule)

result <- foreach(i = 1:ls) %do% {
  set_effects <- combi_plantule[i]
  
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste(resp, "~", effect_formula))
  
  fit <- fitme(formula,
               data = plantule_data,
               family = Gamma(log))
       
  c(extractAIC(fit)[[2]],extractBIC(fit,ntot),set_effects)
}


combiresult <- list()
AICresult <- list()
BICresult <- list()
for(i in 1:length(result)){ 
  AICresult[i] <- result[[i]][1]
  BICresult[i] <- result[[i]][2]
  combiresult[i] <- result[[i]][3]
}


posAIC <- order(unlist(AICresult))[1:5]
posBIC <- order(unlist(BICresult))[1:5]

pltAIC <- AICresult[posAIC]
pltBIC <- BICresult[posBIC]
pltmodAIC <- combiresult[posAIC]
pltmodBIC <- combiresult[posBIC]

pltAIC
pltmodAIC

pltBIC
pltmodBIC 
```
## Establishment rate 

Remplir les données manquantes de nombres de capitules avec des prédictions.
```{r}
cptl_data <- IPM_data[!IPM_data$Flowering==0,]
Cptlglm1 <- fitme(log(Capitule) ~ 1 + Size0Mars + (Age|year), 
              data=cptl_data)
# NbrCptl = exp(2.32734+0.06752*Size0Mars)

cptl_data_predi <- cptl_data %>% 
  mutate(Capitule = ifelse(is.na(Capitule), exp(2.32734+0.06752*Size0Mars), Capitule))
```

```{r}
plt <- IPM_data %>% 
  filter(Age==1) %>% 
  group_by(Quadrat,year,Pop) %>% 
  summarize(NombrePlantules = sum(Age))
  
cptl <- cptl_data_predi %>% 
  group_by(Quadrat,year,Pop) %>% 
  summarize(Capitule = sum(Capitule))
```

### En utilisant le taux d'établissement comme variable
```{r}
Estb <- full_join(plt,cptl, by=join_by(Quadrat,year,Pop))
# summary(Estb)

Estb <- Estb %>% mutate(EstbRate=rep(NA)) %>% 
  arrange(Quadrat)

for (i in 2:length(Estb$Quadrat)){
  if (Estb$Quadrat[i]!=Estb$Quadrat[i-1]){next}
  if (Estb$year[i]!=Estb$year[i-1]+1){next}
  Estb$EstbRate[i] <- Estb$NombrePlantules[i]/Estb$Capitule[i-1]
}
summary(Estb)
```

```{r}
resp <- "EstbRate"
n <- length(Estb$EstbRate)
ls <- length(combi_plantule)

result <- foreach(i = 1:ls) %do% {
  set_effects <- combi_plantule[i]
  
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste(resp, "~", effect_formula))
  
  fit <- fitme(formula,
               data = Estb)
  
  extractAIC(fit)
}


pos <- order(unlist(result))[1:5]

estbAIC <- result[pos]
estbmod <- combi_plantule[pos]

estbAIC
estbmod
```


### En utilisant le nombre de plantules comme variable

```{r}
Estb <- left_join(plt, cptl, by = join_by(Quadrat, year, Pop))
```

```{r}
resp <- "NombrePlantules"
n <- length(Estb$year)
ls <- length(combi_estb)

result <- foreach(i = 1:ls) %do% {
  set_effects <- combi_estb[i]
  
  effect_formula <- paste(unlist(set_effects), collapse = "+")
  formula <- as.formula(paste(resp, "~", effect_formula))
  
  fit <- fitme(formula,
               data = Estb,
               family = Poisson(log))
  
  extractAIC(fit)[[2]]
}


pos <- order(unlist(result))[1:5]

estbAIC <- result[pos]
estbmod <- combi_estb[pos]

estbAIC
estbmod
```
