---
pdf_document: default
author: "Loïc Pages"
date: "`r Sys.Date()`"
output:
  html_document: null
  df_print: paged
  pdf_document: default
word_document: default
title: "IPM results (mean kernel)"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/media/loic/Commun/0Travail/2025 IPM Centauree ISEM/2025 Stage ISEM/Code")
library(spaMM)
library(tidyverse)
library(popbio)
library(splines)
```

# Initialisation

```{r}
GrowthRateMPM <- read.csv("MPM/Growth.rates.csv",sep = " ",dec = ",")
IPM_data <- read.csv("Data/newdata.csv")

MatrixDim <- 50

#Forcer l'age maximal à 8
AgeMax <- 8
IPM_data$Age[IPM_data$Age > AgeMax] <- AgeMax

centauree_data <- IPM_data[!is.na(IPM_data$Size0Mars) & !is.na(IPM_data$Age),]

spaMM.options(separation_max=70)
```

## Vital rate functions

Load selected models

```{r}
load("IPM/ModelsAIC.RData")
```

Creation of the functions for the kernels.


Creation of the functions for the kernels. 
```{r}
annees <- 1995:2022
populations <- c("Po","Au","Pe","E1","E2","Cr")
#ysup=yI.a[,2]; yinf=yI.a[,1]; x=ymid.a; a=2

#Fecundity - Number of capitulas
numbofcap <- function(x, a) {
  fake_data <- expand.grid( #Create a false data set with all combinations of variables
  year = annees,
  Pop = populations,
  Size0Mars = x,
  Age = a) 
  predi <- exp(predict(Cptlglm1, fake_data)) #Predicted number of capitulas
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) #Make the prediction data more usable, and take the mean over the populations and years for each size/age class.
  
  # meanpredi$V1[meanpredi$Group.1<5] <- 0 #Nbr of capitulas=0 if size<5 cm
  nbcap <- ifelse(meanpredi$V1>=0,meanpredi$V1,0) #Avoid nbr of capitulas=0
  return(matrix(nbcap, length(x), length(x), byrow = T))
}

#Survival Probability for age 1
sx01 <- function(x,a) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = x,
  Age = a)
  predi <- predict(Survglm11, fake_data, allow.new.levels = T)
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(meanpredi$V1)
}

#Survival Probability for age > 1
sx02 <- function(x, a) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = x,
  Age = a)
  predi <- predict(Survglm12, fake_data, allow.new.levels = T)
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(meanpredi$V1)
}


#Flowering Probability

#Flowering function for The survival-growth kernel
flr0 <- function(x, a, Beta=0) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  predi <- plogis(predict(Flowglm1, newdata = fake_data, allow.new.levels = T, type = "link") + Beta) 
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(meanpredi$V1)
}

#Flowering function for fecundity fyx0 - same function the difference is in the format of the output
flr1 <- function(x, a, Beta=0) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  mu <- plogis(predict(Flowglm1, fake_data, allow.new.levels = T, type = "link") + Beta)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(matrix(mu2$V1, MatrixDim, MatrixDim, byrow = T))
}

# Growth function (using cumulative distribution function)
#Gaussian
Gyx1 <- function(ysup,yinf,x,a){  
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  size1predi <- predict(Growthglm1, fake_data, allow.new.levels = T) #Prédit les taille à t+1 possibles 
  mu <- aggregate(size1predi, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  var1predi <- sqrt(exp(1.1013346+0.5553254*log(unique(x))-0.1231723*log(a))) #Prédit standard deviation 
  # A modifier si le modèle change, voir phi

  cdf.sup <- mapply(pnorm, mean = mu$V1, sd = var1predi, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = mu$V1, sd = var1predi, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
}

## Gamma(log)
Gyx2 <- function(ysup,yinf,x,a){
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  predi <- predict(Growthglm1gamma, fake_data, allow.new.levels = T) #Prédit les taille à t+1 possibles 
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) 
  phi <- exp(-1.07986 - 0.57167*log(unique(x)) + 0.06724*log(a))
  varpredi <- phi*meanpredi$V1

  cdf.sup <- mapply(pnorm, mean = meanpredi$V1, sd = varpredi, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = meanpredi$V1, sd = varpredi, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
}

#Transformed Gaussian
Gyx3 <- function(ysup,yinf,x,a){  
  sizepow <- 0.4343354
  yinfpow <- yinf**sizepow
  ysuppow <- ysup**sizepow
  yinfpow[1] <- -Inf
  ysuppow[length(ysuppow)] <- Inf
  
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  predi <- predict(Growthglm2, fake_data, allow.new.levels = T) #Predicted size^pow at t+1 for all combinations of pop, year, age and size at t
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) #Make the prediction data more usable, and take the mean over the populations and years for each size/age class.
  
  resvar <- predict(Growthglm2$resid_fit, newdata = fake_data, allow.new.levels = T)
  meanresvar <- aggregate(resvar[,1], list(fake_data$Size0Mars), mean)$x
  
  cdf.sup <- mapply(pnorm, mean = meanpredi$V1, sd = sqrt(meanresvar), MoreArgs = list(q=ysuppow))
  cdf.inf <- mapply(pnorm, mean = meanpredi$V1, sd = sqrt(meanresvar), MoreArgs = list(q=yinfpow))
  return(cdf.sup - cdf.inf)
}

#Seedling size (using cumulative distribution function)
densSeedl <- function(ysup, yinf) {
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  fake_data <- expand.grid(
  year = annees,
  Pop = populations)
  
  predi <- predict(Pltglm1, newdata = fake_data)[, 1]
  meanpredi <- mean(predi)
  shape <- 1 / Pltglm1$phi
  scale <- meanpredi * Pltglm1$phi
  cdf.sup <- pgamma(ysup, shape = shape, scale = scale)
  cdf.inf <- pgamma(yinf, shape = shape, scale = scale)
  return(cdf.sup - cdf.inf)
}

#Establishment rate
Estbli <- function(Pop=populations, year=annees){
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Capitule = 1:200)
  plt <- predict(Estbglm2, newdata = fake_data)[,1]
  mu <- mean(plt/fake_data$Capitule)
  return(mu)
}

#Combine fecundity and flowering probability 
fyx0 <- function(x , a, Estbl, Beta = 0) {
  p.flow <- flr1(x, a, Beta) #flowering probability
  n.captl <- numbofcap(x, a) #number of capitula per flowering plant
  Estbl <- Estbli(populations,annees)
  
  feco <- p.flow * n.captl * Estbl
  return(feco)
}
```

# Kernel

```{r}
load("IPM/Kernals implementation/IPMKernal_AIC.RData")

print(lambda(IPMKernal_AIC))
```

Affichage du kernel général et de quelques sous-matrices

```{r}
M1 <- IPMKernal_AIC[51:100,1:50]
M2 <- IPMKernal_AIC[101:150,51:100]
M3 <- IPMKernal_AIC[151:200,101:150]
M8 <- IPMKernal_AIC[351:400,301:350]

eps <- 1e-4
for (i in 1:50){
  for (j in 1:50){
    if(M1[i,j]<eps){M1[i,j] <- 0}
    if(M2[i,j]<eps){M2[i,j] <- 0}
    if(M3[i,j]<eps){M3[i,j] <- 0}
    if(M8[i,j]<eps){M8[i,j] <- 0}
  }
}
image2(IPMKernal_AIC,cex=0.1)
image2(M2,cex=0.3)
image2(M3,cex=0.3)
image2(M8,cex=0.3)
```

# Fonction de croissance

```{r}
n.size <- 50
G <- list()
df_growth <- list()

minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b <- minsize + c(0:n.size) * h

#sizes for an individual of age a
#midpoint
ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
#intervals
yI.a <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size) {
  yI.a[i, 1] <- b[i]
  yI.a[i, 2] <- b[i + 1]
}

b1 = minsize + c(0:n.size) * h1
#sizes for an individual of age 1
#midpoint
ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
#intervals
yI.1 <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size) {
  yI.1[i, 1] <- b1[i]
  yI.1[i, 2] <- b1[i + 1]
}

for(age in 1:8){
  if (age == 1) {x <- ymid.1} else {x <- ymid.a}
  G <- Gyx3(ysup = yI.a[, 2],yinf = yI.a[, 1], x = x ,a = age)
  
  df_age <- as.data.frame(G) %>%
    mutate(size_t1 = x) %>%
    pivot_longer(cols = -size_t1,
               names_to = "size_t",
               values_to = "prob") %>%
    mutate(size_t = ymid.a[as.numeric(gsub("V", "", size_t))], age = age)

    df_growth[[age]] <- df_age
}
df_growth_all <- bind_rows(df_growth)

```

```{r,echo=FALSE}
maxprob <-  max(df_growth_all$prob)

ggplot(df_growth[[1]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 1",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[2]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 2",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[3]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 3",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[4]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 4",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[5]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 5",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[6]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 6",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[7]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 7",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

ggplot(df_growth[[8]], aes(x = size_t, y = size_t1, fill = prob)) +
  geom_tile() +
  scale_fill_gradientn(
    colours = c("darkblue", "purple", "red", "orange", "yellow"),
    name = "Probabilité",
    limits = c(0, maxprob)
  ) +
  labs(title = "age = 8",
    x = "Taille à t",
    y = "Taille à t+1"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )
```


# Results

## Elasticity

```{r}
library(IPMpack)
library(viridis)

# Paramètres
n_size <- 50
n_age <- 8

K.sens <- IPMpack::sens(IPMKernal_AIC)
K.elas <- IPMpack::elas(IPMKernal_AIC)

eps <- 1e-6
for (i in 1:400){
  for (j in 1:400){
    if(K.elas[i,j]<eps){K.elas[i,j] <- 0}
  }
}
```

```{r}
color_palette <- rev(heat.colors(100))
layout(matrix(c(1, 2), nrow = 1), widths = c(4, 1))

## Matrice d'élasticité
par(mar = c(4, 4, 2, 1))  # marges normales à gauche
image2(K.elas, col = color_palette, cex = 0.01)

## Légende
par(mar = c(4, 2, 2, 4))  # marges pour la colonne légende
image(
  x = 1,
  y = seq(0, 1, length.out = 100),
  z = t(matrix(1:100, ncol = 1)),  # transpose pour verticalité
  col = color_palette,
  axes = FALSE,
  xlab = "",
  ylab = ""
)

# Axe des valeurs
axis(
  side = 4,
  at = seq(0, 1, length.out = 5),
  labels = round(seq(0, max(K.elas, na.rm = TRUE), length.out = 5), 3),
  las = 1
)
mtext("Elasticity", side = 4, line = 2.5)
```

## Stable stage distribution

### Stable size distribution

```{r}
A <- eigen.analysis(IPMKernal_AIC)

n.size <- 50
minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b <- minsize + c(0:n.size) * h
b1 <- minsize + c(0:n.size) * h1

ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
ymid.1 = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])

AA <- tibble(
  Size1 = rep(ymid.1,8),
  Size = rep(ymid.a,8),
  density = rep(NA,50*8),
  age = rep(NA,50*8)
)
  
for (i in 1:8){
  AA$density[(50*(i-1)+1):(50*i)] <- A$stable.stage[(50*(i-1)+1):(50*i)]
  AA$age[(50*(i-1)+1):(50*i)]  <- rep(i,50)
  }
```

Histogramme de tailles observées

```{r}
hist_data_list <- list()

# Boucle sur les âges
for (a in 1:8) {
  data_age <- centauree_data %>% filter(Age == a)
  if (a == 1) {
    bins <- b1
    mids <- ymid.1
  } else {
    bins <- b
    mids <- ymid.a
  }
  hist_counts <- hist(data_age$Size0Mars, breaks = bins, plot = FALSE)
  df_hist <- data.frame(
    Age = a,
    x = mids,
    count = hist_counts$counts
  )
  hist_data_list[[a]] <- df_hist
}

# Combine tout
hist_data <- bind_rows(hist_data_list)

# Normalisation
N_total <- sum(hist_data$count)
hist_data <- hist_data %>%
  mutate(prob = count / N_total,
         age=Age)
```

```{r}
AA %>% 
  ggplot()+
  geom_col(data=hist_data,aes(x = x, y = prob), fill = "steelblue", color = "black", width = h) +
  geom_line(aes(x=Size,y=density),col="red")+
  # geom_smooth(aes(x=Size,y=density))+
  theme_bw()+
  facet_wrap(~age,scale="free")+
  labs(x="Size",
      y="Density")
```

### Stable age distribution

```{r}
age_dist <- hist_data %>%
  group_by(age) %>% 
  summarize(prob = sum(prob))

AA %>% group_by(age) %>% 
  summarize(density = sum(density)) %>% 
  ggplot()+
  geom_col(data=age_dist, aes(x = age, y = prob, fill = as.factor(age)), color = "black", width = 0.5, show.legend = FALSE) +
  scale_fill_brewer(palette = "Spectral", direction = -1) +
  geom_line(aes(y=density,x=age),col="red",linewidth=1)+
  theme_bw()+
  labs(x="Age",
       y="Density")

```

### Stable size and age at flowering

```{r}
summary(IPM_data$Age[IPM_data$Flowering==1])
af <- hist(
  IPM_data$Age[IPM_data$Flowering == 1],
  breaks = seq(0.5, 8.5, by = 1),
  xaxt = "n",
  xlab = "Age",
  ylab = "Number of flowering plants",
  main = "")
axis(1, at = 1:8, labels = 1:8)
```

```{r}
A <- eigen.analysis(IPMKernal_AIC)
N <- matrix(ncol = 8, nrow = 50)
for (i in 1:8){
  N[,i] <- A$stable.stage[(50*(i-1)+1):(50*i)]
}
W <- function(age){
  return(N[(50*(age-1)+1):(50*age)])
}
```

```{r}
minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
n.size <- MatrixDim
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b = minsize + c(0:n.size) * h

#sizes for an indiWidual of age a
#midpoint
ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
#interWals
yI.a <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.a[i,1] <- b[i]
  yI.a[i,2] <- b[i+1]
}

b1 = minsize + c(0:n.size) * h1
#sizes for an indiWidual of age 1
#midpoint
ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
#interWals
yI.1 <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.1[i,1] <- b1[i]
  yI.1[i,2] <- b1[i+1]
}


```

Distribution de taille des plantes en fleurs

```{r}
#Proportion de plantes en fleurs dans la population
Flr_dens <- sum(flr0(ymid.1,1)*W(1))
for (a in 2:8) { 
  Flr_dens <- Flr_dens + sum(flr0(ymid.a, a) * W(a))
}

Flry <- rep(NA, 50) #Distribution de taille des plantes en fleur
Z <- rep(NA, 50)
for (i in 1:MatrixDim) {
  Flry[i] <- sum(flr0(ymid.1[i],1)*W(1)[i])
  for (a in 2:8) {
    Flry[i] <- Flry[i] + flr0(ymid.a[i], a) * W(a)[i]
  }
}
Z <- Flry/Flr_dens #normalisation

flower_data <- centauree_data[centauree_data$Flowering==1 & !is.na(centauree_data$Flowering),]
flower_data <- flower_data %>%
  group_by(Size0Mars) %>%
  summarise(n_fleuri = n()) %>%
  mutate(Wobs = n_fleuri / sum(n_fleuri),
         Size=Size0Mars) %>% dplyr::select(!n_fleuri) %>% 
  dplyr::select(!Size0Mars)

tibble(Size = seq(0.5,25,0.5),
                 West = Z) %>% 
  left_join(flower_data,join_by(Size)) %>% 

  ggplot(aes(x = Size)) +
  geom_bar(aes(y = Wobs), stat = "identity", fill = "steelblue",color="black") +
  geom_line(aes(y = West), color = "red", linewidth = 1) +
  labs(x = "Size at flowering", y = "Density")+
  theme_bw()+
      theme(
    axis.title.x = element_text(size = 16),  # Taille du titre axe X
    axis.title.y = element_text(size = 16),  # Taille du titre axe Y
    axis.text.x = element_text(size = 14),   # Taille des graduations axe X
    axis.text.y = element_text(size = 14)    # Taille des graduations axe Y
  )
```


# Calcul des traits d'histoire de vie correspondant au modèle matriciel

```{r}
Fec <- function(i,a){ 
  #i indice de taille de plantule 
  #a age de plante mère
  return(IPMKernal_AIC[i,((a-1)*50+1):(a*50)])
}

P <- function(i,a){
  #i indice de taille de plante à t+1 
  #a age de plante à t
  if(a==8){return(IPMKernal_AIC[((50*(a-1))+i),((a-1)*50+1):(a*50)])}
  return(IPMKernal_AIC[((50*a)+i),((a-1)*50+1):(a*50)])
}
```

N1: densité de plantules N2: densité de rosettes N3: densité d'individus en fleurs $$
    fs_0=\frac{\sum\limits_{a=1}^8\sum\limits_i\sum\limits_j W_{a,j}\tilde{F}(j,a,i,1)}{\sum\limits_{a=1}^8\sum\limits_jW_{a,j}f(j,a)}
$$

```{r}
N1t1 <- 0
N3t0 <- sum(W(1)*flr0(ymid.1,1))
for(a in 2:8){
  N3t0 <- N3t0 + sum(W(a)*flr0(ymid.a,a))
  }
for(a in 1:8){
  for (i in 1:MatrixDim){
    N1t1 <- N1t1 + sum(Fec(i,a) * W(a))
  }
}
fs0 <- N1t1/N3t0
fs0
```

$$   s_1(1-\alpha_1)=\frac{\sum\limits_i\sum\limits_j W_{1,j}\tilde{P}(j,1,i,2)(1-f(i,2))}{\sum\limits_jW_{1,j}(1-f(j,1))}$$

```{r}
N2t1 <- 0
N1t0 <- sum(W(1)*(1-flr0(ymid.1,1)))
for (i in 1:MatrixDim){
  N2t1 <- N2t1 + sum(P(i,1) * W(1) * (1-flr0(ymid.a[i],2)))
}
s1_1a1 <- N2t1/N1t0
s1_1a1
```

$$ s_2(1-\alpha_2)=\frac{\sum\limits_i\sum\limits_j \sum\limits_{a=2}^7W_{a,j}\tilde{P}(j,a,i,a+1)(1-f(i,a+1))+ W_{8,j}\tilde{P}(j,8,i,8)(1-f(i,8))}{\sum\limits_{a=2}^8\sum\limits_jW_{a,j}(1-f(j,a))}$$

```{r}
N2t1 <- 0
N2t0 <- 0
for(a in 2:8){
  N2t0 <- N2t0 + sum(W(a)*(1-flr0(ymid.a,a)))
}
for (i in 1:MatrixDim){
  for(a in 2:7){
      N2t1 <- N2t1 + sum(P(i,a) * W(a) * (1-flr0(ymid.a[i],a+1)))
  }
  N2t1 <- N2t1 + sum(P(i,8) * W(8) * (1-flr0(ymid.a[i],8)))
}
s2_1a2 <- N2t1/N2t0
s2_1a2
```

$$s_2\alpha_2=\frac{\sum\limits_i\sum\limits_j \sum\limits_{a=2}^7W_{a,j}\tilde{P}(j,a,i,a+1)f(i,a+1)+ W_{8,j}\tilde{P}(j,8,i,8)f(i,8)}{\sum\limits_{a=2}^8\sum\limits_jW_{a,j}(1-f(j,a))}$$

```{r}
N2t1 <- 0
N2t0 <- 0
for(a in 2:8){
  N2t0 <- N2t0 + sum(W(a)*(1-flr0(ymid.a,a)))
}
for (i in 1:MatrixDim){
  for(a in 2:7){
      N2t1 <- N2t1 + sum(P(i,a) * W(a) * flr0(ymid.a[i],a+1))
  }
  N2t1 <- N2t1 + sum(P(i,8) * W(8) * flr0(ymid.a[i],8))
}
s2_a2 <- N2t1/N2t0
s2_a2
```

```{r}
IPM <- matrix(c(0,s1_1a1,0,0,s2_1a2,s2_a2,fs0,0,0),nrow=3,ncol=3)

load("MPM/matrice_globale.RData")
MPM <- A
rm(A)

IPM
MPM
```

Taille et age observés moyen à la floraison

```{r}
mu_size <- sum(ymid.1*W(1)*flr0(ymid.1,1))
mu_a <- sum(W(1)*flr0(ymid.1,1))
mu2 <- sum(W(1)*flr0(ymid.1,1))
for(a in 2:8){
  mu_size <- mu_size + sum(ymid.a*W(a)*flr0(ymid.a,a))
  mu_a <- mu_a + sum(a*W(a)*flr0(ymid.a,a))
  mu2 <- mu2 + sum(W(a)*flr0(ymid.a,a))
}
mu_size <- mu_size/mu2
mu_a <- mu_a/mu2
mu_size
mu_a
```

