---
title: "Integrated Population Model"
author: "Eric Imbert & Loïc Pages"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

# Initialisation
```{r}
rm(list=ls())
library(knitr)
library(spaMM)
library(tidyverse)
library(popbio)
library(splines)

setwd("/media/loic/Commun/0Travail/Stage 2025 ISEM/Code/IPM")

GrowthRateMPM <- read.csv("Growth.rates.csv",sep = " ",dec = ",")
IPM_data <- read.csv("newdata.csv")

MatrixDim <- 50
RecrRate <- 0.338 #recruitment rate

#Forcer l'age maximal à 8
AgeMax <- 8
IPM_data$Age[IPM_data$Age > AgeMax] <- AgeMax

centauree_data <- IPM_data[!is.na(IPM_data$Size0Mars) & !is.na(IPM_data$Age),]

spaMM.options(separation_max=70)
```

Age of flowering plants - will be used later to compute optimal growth rate
```{r}
weightAgeFlrPool=NULL
for ( a in 1:AgeMax ) {
  weightAgeFlrPool[a]= nrow(centauree_data[which(centauree_data$Age==a & centauree_data$Flowering==1),])
}
```

Seedling size distribution
```{r}
plantule_data <- IPM_data[IPM_data$Age==1,] %>% 
  filter(!is.na(Age)) %>% 
  filter(Size0Mars!=0)

n_plt <- length(plantule_data$Individu)

sizeSeedl <- unique(plantule_data$Size0Mars[order(plantule_data$Size0Mars)])

Width <-NULL 
Den <- NULL 

for (i in 1:length(sizeSeedl)){
  # Number of observations for each size
  Width[i] <- length(unique(plantule_data$Individu[plantule_data$Size0Mars==sizeSeedl[i]]))
  # Frequency of each size
  Den[i] <- Width[i]/n_plt
}
Width
Den
dens <- rep(Den,Width)

gg1=lm(log(1/dens)~rep(sizeSeedl,Width))

densSeedl2=function(y,intervalle){   
  miny=min(y)
  maxy=max(y)
  SeedlgSize=exp(log(coef(gg1)[2])-y*coef(gg1)[2])/(exp(-coef(gg1)[2]*(miny-intervalle/2))-exp(-coef(gg1)[2]*(maxy+intervalle/2)))
  return(SeedlgSize)
}
```

## Models
Create a false data set 
```{r}
## Creation of fake data
AgeMax <- 8
year <- 1995:2022
Pop <- c("Po","Au","Pe","E1","E2","Cr")
taille_range <- seq(0.5, 25, by = 0.5) 
age_range <- 1:AgeMax
# 
# fake_data <- expand.grid(
#   year = annees,
#   Pop = populations,
#   Size0Mars = NA,
#   Age = NA)


Pop=gl(6,length(year),labels = Pop)
rer=data.frame(Pop,year)
rer$Index=1:(length(year)*6)
fake_data=do.call("rbind", replicate(MatrixDim, rer, simplify = FALSE))
fake_data=fake_data[order(fake_data$Index),]
fake_data$Size0Mars=NA
fake_data$Age=NA

rm(rer, Pop, year)
```

Load the selected models
```{r}
load("ModelsAIC")
```

Creation of the functions for the kernels. 
```{r}
annees <- 1995:2022
populations <- c("Po","Au","Pe","E1","E2","Cr")

#Fecundity
numbofcap <- function(x, a) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = x,
  Age = a)
  mu <- exp(predict(Cptlglm1, fake_data))
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  nbcap <- ifelse(mu2$V1>=0,mu2$V1,0)
  return(matrix(nbcap, length(x), length(x), byrow = T))
}

#Survival Probability for age 1
sx01 <- function(x,a) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = x,
  Age = a)
  mu <- predict(Survglm11, fake_data, allow.new.levels = T)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}

#Survival Probability for age > 1
sx02 <- function(x, a) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = x,
  Age = a)
  mu <- predict(Survglm12, fake_data, allow.new.levels = T)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}


#Flowering Probability
#Beta = will be used to modify the intercept of the flowering function
#obs_beta = observed value
#extract Beta0
load("obs_beta")
load("se_obs_beta")


#Flowering function for The survival-growth kernel
flr0 <- function(x, a, Beta) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  mu <- plogis(predict(Flowglm1, newdata = fake_data, allow.new.levels = T, type = "link") + Beta)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}

#Flowering function for fecundity fyx0 - same function the difference is in the format of the output
flr1 <- function(x, a, Beta) {
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  mu <- plogis(predict(Flowglm1, fake_data, allow.new.levels = T, type = "link") + Beta)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(matrix(mu2$V1, MatrixDim, MatrixDim, byrow = T))
}

# Méthode cumulative ditribution function
Gyx2 <- function(ysup,yinf,x,a){  
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = annees,
  Pop = populations,
  Size0Mars = unique(x),
  Age = a)
  mu <- predict(Growthglm2, fake_data, allow.new.levels = T) #Prédit les taille à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  usd <- sqrt(exp(1.1013346+0.5553254*log(unique(x))-0.1231723*log(a))) #Prédit standard deviation 
  # A modifier si le modèle change, voir phi

  cdf.sup <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
  }

#Seedling size density
densSeedl <- function(ysup, yinf) {
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  mu <- predict(Pltglm1, newdata = fake_data)[, 1]
  shape <- 1 / Pltglm1$phi
  mu <- mean(mu)
  scale <- mu * Pltglm1$phi
  cdf.sup <- pgamma(ysup, shape = shape, scale = scale)
  cdf.inf <- pgamma(yinf, shape = shape, scale = scale)
  return(cdf.sup - cdf.inf)
}

Estbli <- function(Pop=populations, year=annees){
  fake_data <- expand.grid(
  year = year,
  Pop = Pop)
  mu <- mean(predict(Estbglm1, newdata= fake_data))
  # sample(predict(Estbglm1),1)
  return(mu)
}

#Combine fecundity and flowering probability 
fyx02 <- function(x , a, Estbl, Beta = 0) {
  p.flow <- flr1(x, a, Beta) #flowering probability
  n.captl <- numbofcap(x, a) #number of capitula per flowering plant
  
  Fyx0 <- p.flow * n.captl * Estbl 
  return(Fyx0)
}
```

```{r}
n.size <- 50

K.fnc0 <- function(age, n.size, estbl, Beta=0){
  
  minsize <-  0.5
  maxsize <-  25
  maxsize1 <- 15
  h <-  (maxsize - minsize) / n.size
  h1 <-  (maxsize1 - minsize) / n.size
  b <- minsize + c(0:n.size) * h

  #sizes for an individual of age a
  #midpoint
  ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
  #intervals
  yI.a <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.a[i,1] <- b[i]
    yI.a[i,2] <- b[i+1]
  }
  
  b1 = minsize + c(0:n.size) * h1
  #sizes for an individual of age 1
  #midpoint
  ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
  #intervals
  yI.1 <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.1[i,1] <- b1[i]
    yI.1[i,2] <- b1[i+1]
  }
  
  if(age==1){h <- h1
    ymid.a <- ymid.1
  }

  #The growth kernel
  G <- Gyx2(ysup=yI.a[,2], yinf=yI.a[,1], x=ymid.a, a=age)

  #The survival-growth kernel #remove flowering plants since they died
  sx0 <- ifelse(age==1,sx01,sx02)
  if(age==1){surv <- sx01(ymid.a,1)}
  if(age!=1){surv <- sx02(ymid.a,age)}
  S_flr <- surv*(1-flr0(ymid.a,age,Beta))

  Ps <- sweep(G, MARGIN=2, S_flr, '*')
  P <- matrix(Ps, nrow=n.size)
  
  #Fecundity kernel
  ProporSeedlSize <- densSeedl(ysup=yI.1[,2],yinf=yI.1[,1])
  feco <- fyx02(x = ymid.a, a = age, Estbl = estbl, Beta = Beta)
  Fec <- sweep(feco, MARGIN=1, ProporSeedlSize, '*')

  IPMk <- P + Fec
  return(list(M=IPMk,Pk=P,Fk=Fec))
    }
```

the super matrix and growth rate
```{r}
super_matrix = function(Beta_test){
  Pk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Fk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Mk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Kernals <- array(data = 0, dim = c(AgeMax, 3, MatrixDim, MatrixDim))
  GlobMatrIPM <- array(0, dim = c((AgeMax * MatrixDim), (AgeMax * MatrixDim)))
  # estbl <- Estbli(Pop,annees)
  
  for (a in (1:(AgeMax - 1))){
    Kls <- K.fnc0(a, MatrixDim, estbl, Beta_test)
    Kernals[a, 1, , ] <- Kls$M
    Kernals[a, 2, , ] <- Kls$Pk
    Kernals[a, 3, , ] <- Kls$Fk
    # print(lambda(Kernals[a, 1, , ]))
    debut <- (a - 1) * MatrixDim + 1
    fin <- debut + (MatrixDim - 1)
    GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
    GlobMatrIPM[((fin + 1):(fin + MatrixDim)), (debut:fin)] <- Kls$Pk
  }
  
  Kls <- K.fnc0(AgeMax, MatrixDim, estbl, Beta_test)
  Kernals[AgeMax, 1, , ] <- Kls$M
  Kernals[AgeMax, 2, , ] <- Kls$Pk
  Kernals[AgeMax, 3, , ] <- Kls$Fk
  # print(lambda(Kernals[a,1,,]))
  debut <- (AgeMax - 1) * MatrixDim + 1
  fin <- debut + (MatrixDim - 1)
  GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
  GlobMatrIPM[(debut:fin), (debut:fin)] <- Kls$Pk
  
  return(GlobMatrIPM)
}
```

```{r}
# estbl <- Estbli(populations,year)

estbl <- RecrRate
Beta <- 0
MatrixIPM <- super_matrix(0) # growthrate values reported in table 1
IPMKernal_AIC <- MatrixIPM
save(IPMKernal_AIC,file="IPMKernal_AIC")

load("IPMKernal_AIC")
MatrixIPM <- IPMKernal_AIC
print(lambda(MatrixIPM))
```

Affichage du kernel général et de quelques sous-matrices
```{r}
M1 <- MatrixIPM[51:100,1:50]
M2 <- MatrixIPM[101:150,51:100]
M3 <- MatrixIPM[151:200,101:150]
M8 <- MatrixIPM[351:400,301:350]

eps <- 1e-4
for (i in 1:50){
  for (j in 1:50){
    if(M1[i,j]<eps){M1[i,j] <- 0}    
    if(M2[i,j]<eps){M2[i,j] <- 0}    
    if(M3[i,j]<eps){M3[i,j] <- 0}
    if(M8[i,j]<eps){M8[i,j] <- 0}

  }
}
image2(MatrixIPM,cex=0.1)
image2(M2,cex=0.3)
image2(M3,cex=0.3)
image2(M8,cex=0.3)
```


```{r}
sous_mat <- function(Matrix, n_subclass, class_row, class_col){
  n_size <- dim(Matrix)[1]/n_subclass
  row <- ((class_row-1)*n_size+1):(class_row*n_size)
  col <- ((class_col-1)*n_size+1):(class_col*n_size)
  sm <- Matrix[row,col]
  return(sm)
}

image2(sous_mat(K,8,4,3),cex=0.1)
```


Elasticity
```{r}
library(IPMpack)
library(viridis)

# Paramètres
n_size <- 50
n_age <- 8

# Plages de tailles
size_range_age1 <- seq(0.5, 15, length.out = n_size)
size_range_other <- seq(0.5, 25, length.out = n_size)

# Chargement de la matrice de sensibilité
K <- MatrixIPM

K.sens <- IPMpack::sens(K)
K.elas <- IPMpack::elas(K)


# Extraire les sous-matrices de sensibilité pour chaque âge
elasticity_list <- list()

for (a in 2:(n_age-1)) {
  # idx <- ((a - 1) * n_size + 1):(a * n_size)
  # submat <- elasticity_mat[idx, idx]
  submat <- sous_mat(K.elas,n_age,a,a)
  # submat <- submat / max(submat)  # Normaliser chaque bloc
  elasticity_list[[a]] <- submat
}
elasticity_list[[1]] <- sous_mat(K.elas,n_age,1,2)
elasticity_list[[8]] <- sous_mat(K.elas,n_age,8,8)

image(size_vector,size_vector,t(K.elas))

eps <- 1e-6
for (i in 1:400){
  for (j in 1:400){
    if(K.elas[i,j]<eps){K.elas[i,j] <- 0}
  }
}
image2(K.elas,cex=0.01)

filled.contour(
  x = size_vector,
  y = size_vector,
  z = t(K.elas),
  col = gray.colors(100,start=1,end=0),
  plot.title = title(main = "Matrice complète de elasticité", 
                     xlab = "Taille à t (z)", 
                     ylab = "Taille à t+1 (z')"),
  key.title = title("Elasticité"),
  plot.axes = {
    axis(1); axis(2)
    contour(size_vector, size_vector, t(K.elas), add = TRUE)
  }
)
# Tracer les sous-matrices avec contours
par(mfrow = c(2, 4), mar = c(4, 4, 2, 1))

for (a in 1:n_age) {
  mat <- elasticity_list[[a]]
  sizes0 <- if (a == 1) size_range_age1 else size_range_other
  
  filled.contour(
    x = sizes0, y = size_range_other, z = t(mat),
    color.palette = gray.colors,
    plot.title = title(main = paste("Age", a), xlab = "Size(t), x", ylab = "Size(t+1), y'"),
    key.title = title("Elasticity"),
    plot.axes = {
      axis(1); axis(2)
      contour(sizes0, size_range_other, t(mat), add = TRUE, drawlabels = TRUE)
    }
  )
}
```

```{r}
color_palette <- rev(heat.colors(100))
# Layout à 2 colonnes : image à gauche, légende à droite
layout(matrix(c(1, 2), nrow = 1), widths = c(4, 1))

## --- 1. Image principale avec image2() ---
par(mar = c(4, 4, 2, 1))  # marges normales à gauche
image2(K.elas, col = color_palette, cex = 0.01)

## --- 2. Barre de légende manuelle ---
par(mar = c(4, 2, 2, 4))  # marges pour la colonne légende
image(
  x = 1,
  y = seq(0, 1, length.out = 100),
  z = t(matrix(1:100, ncol = 1)),  # transpose pour verticalité
  col = color_palette,
  axes = FALSE,
  xlab = "",
  ylab = ""
)

# Axe des valeurs à droite
axis(
  side = 4,
  at = seq(0, 1, length.out = 5),
  labels = round(seq(0, max(K.elas, na.rm = TRUE), length.out = 5), 3),
  las = 1
)
mtext("Elasticity", side = 4, line = 2.5)
```


Stable distribution size by age
```{r}
A <- eigen.analysis(MatrixIPM)
A$lambda1

minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b <- minsize + c(0:n.size) * h
b1 <- minsize + c(0:n.size) * h1
x <- 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
x1 <- 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])

AA <- tibble(
  Size1 = rep(x1,8),
  Size = rep(x,8),
  density = rep(NA,50*8),
  age = rep(NA,50*8)
)
  
for (i in 1:8){
  AA$density[(50*(i-1)+1):(50*i)] <- A$stable.stage[(50*(i-1)+1):(50*i)]
  AA$age[(50*(i-1)+1):(50*i)]  <- rep(i,50)
  }

AA %>% 
  filter(age!=1) %>% 
  ggplot()+
  geom_line(aes(x=Size,y=density,color=as.factor(age)))+
  # geom_smooth(aes(x=Size,y=density,color=as.factor(age)))+
  theme_bw()+
  labs(y = "Density",
       x = "Size",
       color = "Age")
  
AA %>% 
  filter(age==1) %>% 
  ggplot()+
  geom_line(aes(x=Size1,y=density),col="red")+
  theme_bw()+
    labs(y = "Density",
       x = "Size")

```


```{r}
n.size <- 50
minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b <- minsize + c(0:n.size) * h
b1 <- minsize + c(0:n.size) * h1

ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
ymid.1 = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])

hist_data_list <- list()

# Boucle sur les âges
for (a in 1:8) {
  data_age <- centauree_data %>% filter(Age == a)
  if (a == 1) {
    bins <- b1
    mids <- ymid.1
  } else {
    bins <- b
    mids <- ymid.a
  }
  hist_counts <- hist(data_age$Size0Mars, breaks = bins, plot = FALSE)
  df_hist <- data.frame(
    Age = a,
    x = mids,
    count = hist_counts$counts
  )
  hist_data_list[[a]] <- df_hist
}

# Combine tout
hist_data <- bind_rows(hist_data_list)

# Normalisation
N_total <- sum(hist_data$count)
hist_data <- hist_data %>%
  mutate(prob = count / N_total,
         age=Age)


ggplot(hist_data, aes(x = x, y = prob)) +
  geom_col(fill = "steelblue", color = "black", width = h) +
  facet_wrap(~age, scales = "free") +
  labs(
    title = "Histogrammes avec densité normalisée globalement (somme = 1)",
    x = "Size0Mars", y = "Densité normalisée"
  ) +
  theme_minimal()
```


```{r}
AA %>% 
  ggplot()+
  geom_col(data=hist_data,aes(x = x, y = prob), fill = "steelblue", color = "black", width = h) +
  geom_line(aes(x=Size,y=density),col="red")+
  # geom_smooth(aes(x=Size,y=density))+
  theme_bw()+
  facet_wrap(~age,scale="free")+
  labs(x="Size",
      y="Density")
```

Stable age distribution
```{r}
age_dist <- hist_data %>%
  group_by(age) %>% 
  summarize(prob = sum(prob))

AA %>% group_by(age) %>% 
  summarize(density = sum(density)) %>% 
  ggplot()+
  geom_col(data=age_dist, aes(x = age, y = prob, fill = as.factor(age)), color = "black", width = 0.5, show.legend = FALSE) +
  scale_fill_brewer(palette = "Spectral", direction = -1) +
  geom_line(aes(y=density,x=age),col="red",linewidth=1)+
  theme_bw()+
  labs(x="Age",
       y="Density")

```


```{r}
obs_lambda <- super_matrix(0) # growthrate values reported in table 1
print(obs_lambda)
print(obs_beta)

mean_size <- mean(IPM_data$Size0Mars[which(IPM_data$Flowering == 1)])
sd_size <- sd(IPM_data$Size0Mars[which(IPM_data$Flowering == 1)])
print(mean_size)
print(sd_size)

mean_age <- mean(IPM_data$Age[which(IPM_data$Flowering == 1)], na.rm = T)
sd_age <- sd(IPM_data$Age[which(IPM_data$Flowering == 1)], na.rm = T)

print(mean_age)
print(sd_age)
```

test de différentes tailles de matrice (effet sur la précision de "l'intégration")
```{r}
lambk <- NULL
library(foreach)
library(doParallel)

cl <- makeCluster(10)
registerDoParallel(cl)

lambk <- foreach (k=seq(2,50,5),.packages=c("splines","spaMM","popbio")) %dopar% {
  MatrixDim <- k
  obs_lambda <- lambda(super_matrix(0))
  obs_lambda
}
stopCluster(cl)

lambk <- unlist(lambk)
plot(x = (1:length(lambk))*5, y = lambk, type = "l", lwd=2, xlab ="Number of intervals",ylab="Growth rate",cex.lab=1.3)  ; abline(a=lambk[length(lambk)],b=0,col="red",lty=3)
```


age at flowering
```{r}
#Predicted


#Observed
flower_data <- centauree_data[centauree_data$Flowering==1 & !is.na(centauree_data$Flowering),]
flower_data %>% 
  group_by(Pop) %>% 
  mutate(mean.age = mean(Age,na.rm=TRUE)) %>% 
  ggplot(aes(x=Pop, y=Age)) +
  geom_boxplot(fill="grey")+
  geom_point(aes(y=mean.age),color= "red")+
  theme_bw()
```


```{r}
growthdata <- centauree_data[!is.na(centauree_data$Size1Mars), ]
growthdata <- growthdata[growthdata$Size1Mars != 0, ]
growthdata <- growthdata[!is.na(growthdata$Age),]
growthdata %>% 
  group_by(Size0Mars) %>% 
  mutate(observed_sd = sd(Size1Mars),
         predicted_sd = sqrt(exp(1.0462608+0.5367287*log(Size0Mars)))) %>% 
  ggplot(aes(x=Size0Mars))+
  geom_line(aes(y = observed_sd),color="red")+
  geom_line(aes(y = predicted_sd))+
  theme_bw()+
  labs(y = "Residual Variance",
       x = "Size",
       title = "Comparaison entre la variance observée (rouge) 
       et la variance prédite (noire)")
```


Changes in the optimal flowering strategy
```{r}
minbeta <- -25-obs_beta
maxbeta <- 5-obs_beta
beta <- seq(minbeta,maxbeta,1)
lam <- list()
for (i in 1:length(beta)){
  lam[i] <- unlist(super_matrix(beta[i]))
}
lam <- unlist(lam)
pos <- order(lam[1:20])
maxibeta <- beta[pos[length(pos)]]
plot(x=beta+obs_beta,y=lam,type = 'line') ; abline(v=obs_beta,col="green4",lty=2) ; abline(v=obs_beta-se_obs_beta,lty=2) ; abline(v=obs_beta+se_obs_beta,lty=2) ; abline(v=maxibeta+obs_beta,col="red") ; abline(h=0.869,lty=2,col="green4"); abline(h=max(lam[1:20]),col="red",lty=3) 
```


Size at flowering using the stable distribution size at flowering
Method Rees and Rose 2002
```{r,}
B <- 0; mu <- NULL; StablSize <- array(0,dim = c(MatrixDim,AgeMax))
minsize <-  0.5
maxsize <-  max(IPM_data$Size0Mars, na.rm=T)
X <- seq(minsize,maxsize,length=MatrixDim)

for (i in 1:AgeMax) {
  sx0 <- ifelse(i == 1, sx01, sx02)
  Kernals <- K.fnc0(i, MatrixDim, RecrRate, B)
  AA <- eigen.analysis(Kernals$M)
  V <- AA$stable.stage
  # V_y <- function(y){V[y]}

  mm <- lm(log(1/V) ~ X)
  V_y <- function(y) {return(exp(-coef(mm)[1]-coef(mm)[2]*y))}
  Flr <- function(y,a) {return(sx0(y,a)*flr0(y,a,B)*V_y(y) ) }
  int_age <- as.numeric(integrate(Flr, minsize, maxsize, i, stop.on.error = F)$value)
  
  W_y <- function(y) {return(Flr(y,i)/int_age)}
  Z_y <- function(y,a) {return(W_y(y)*y)}

  Mu <- integrate(Z_y, minsize, maxsize, i, stop.on.error = F) #Mean size at flowering
  mu[i] <- (Mu$value) #Mean size at flowering
  StablSize[, i] <- 100 * (W_y((X)))
}


S <- rowSums(StablSize)
W <- weightAgeFlrPool/sum(weightAgeFlrPool )
mu_glob <- sum(mu*W) #Predicted Mean size at flowering
mu_glob

observed_mean_value <- mean(centauree_data$Size0Mars[which( centauree_data$Age!="NA" & centauree_data$Flowering==1)])
observed_mean_value
observed_sd_value <- sd(centauree_data$Size0Mars[which( centauree_data$Age!="NA" & centauree_data$Flowering==1)])
observed_sd_value


hist(centauree_data$Size0Mars[which( centauree_data$Age!="NA" & centauree_data$Flowering==1 )],xlim = c(0,30),col="blue4",main = "",xlab = "Size of flowering plants (cm)",  ylab = "Nbr. of flowering individuals",   freq = T ,ylim= c(0, 80),breaks =10);lines(X,S,lwd=2,col="green4")
```

```{r}
barplot(AA$stable.stage, col="green", ylim=c(0,0.05), ylab="Stable stage proportion", xlab="Stage class", main="Teasel")
image2(AA$elasticities, cex=.1, mar=c(0.5,3,4,1) )
image2(AA$sensitivities, cex=.1, mar=c(0.5,3,4,1) )
```
