---
  title: "Comparaison between IPM prediction and MPM Matrix"
author: "Eric Imbert & Loïc Pages"
date: "`r Sys.Date()`"
output:
  html_document:
  df_print: paged
word_document: default
pdf_document: default
---
  
# Initialisation
```{r}
rm(list=ls())
library(knitr)
library(spaMM)
library(tidyverse)
library(popbio)
library(splines)

setwd("/media/loic/Commun/0Travail/Stage 2025 ISEM/Models/IPM")

centauree_data <- read.csv("donnesIPM_short.csv")
centauree_data_complet <- read.csv("donnesIPM.csv")

MatrixDim <- 50
RecrRate <- 0.407 #recruitment rate

#Forcer l'age maximal à 8
AgeMax <- 8
centauree_data$age1 <- ifelse(centauree_data$Stage1=="V",centauree_data$age0+1,NA)
centauree_data$age0[centauree_data$age0 > AgeMax] <- AgeMax
centauree_data$age1[centauree_data$age1 > AgeMax] <- AgeMax

spaMM.options(separation_max=70)
```

Age of flowering plants - will be used later to compute optimal growth rate
```{r}
weightAgeFlrPool=NULL
for ( a in 1:AgeMax ) {
  weightAgeFlrPool[a]= nrow(centauree_data[which(centauree_data$age0==a & centauree_data$Flowering0==1),])
}
```

Seedling size distribution
```{r}
plantule_data <- centauree_data_complet[centauree_data_complet$age0==1,] %>% 
  filter(!is.na(age0)) %>% 
  filter(Size0Mars!=0)

n_plt <- length(plantule_data$Nrw)

sizeSeedl <- unique(plantule_data$Size0Mars[order(plantule_data$Size0Mars)])

Width <-NULL 
Den <- NULL 

for (i in 1:length(sizeSeedl)){
  # Number of observations for each size
  Width[i] <- length(unique(plantule_data$Nrw[plantule_data$Size0Mars==sizeSeedl[i]]))
  # Frequency of each size
  Den[i] <- Width[i]/n_plt
}
Width
Den
dens <- rep(Den,Width)

gg1=lm(log(1/dens)~rep(sizeSeedl,Width))

densSeedl2=function(y,intervalle){   
  miny=min(y)
  maxy=max(y)
  SeedlgSize=exp(log(coef(gg1)[2])-y*coef(gg1)[2])/(exp(-coef(gg1)[2]*(miny-intervalle/2))-exp(-coef(gg1)[2]*(maxy+intervalle/2)))
  return(SeedlgSize)
}
```
# Global

## Models

Load the false data set which contains the predicted responses for survival probability, flowering probability, number of capitules (fecondity) and growth. There is a value of each of these response for each combination of size, age, population and year.
```{r}
## Creation of fake data
AgeMax <- 8
year <- 1995:2022
Pop <- c("Po","Au","Pe","E1","E2","Cr")
taille_range <- seq(0.5, 25, by = 0.5) 
age_range <- 1:AgeMax
# 
# fake_data <- expand.grid(
#   year = annees,
#   Pop = populations,
#   Size0Mars = NA,
#   age0 = NA)


Pop=gl(6,length(year),labels = Pop)
rer=data.frame(Pop,year)
rer$Index=1:(length(year)*6)
fake_data=do.call("rbind", replicate(MatrixDim, rer, simplify = FALSE))
fake_data=fake_data[order(fake_data$Index),]
fake_data$Size0Mars=NA
fake_data$age0=NA

rm(rer, Pop, year)

load("ModelsAIC")

growthdata <- centauree_data[!is.na(centauree_data$Size1Mars), ]
growthdata <- growthdata[growthdata$Size1Mars != 0, ]
growthdata <- growthdata[!is.na(growthdata$age0),]
```

Creation of the functions for the kernels. 
```{r}
annees <- 1995:2022
populations <- c("Po","Au","Pe","E1","E2","Cr")

#Fecundity
numbofcap <- function(x, a) {
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = x,
    age0 = a)
  mu <- exp(predict(Cptlglm1, fake_data))
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  nbcap <- ifelse(mu2$V1>=0,mu2$V1,0)
  return(matrix(nbcap, length(x), length(x), byrow = T))
}

#Survival Probability for age 1
sx01 <- function(x,a) {
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = x,
    age0 = a)
  mu <- predict(Survglm11, fake_data, allow.new.levels = T)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}

#Survival Probability for age > 1
sx02 <- function(x, a) {
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = x,
    age0 = a)
  mu <- predict(Survglm12, fake_data, allow.new.levels = T)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}


#Flowering Probability
#Beta = will be used to modify the intercept of the flowering function
#obs_beta = observed value
#extract Beta0
load("obs_beta")
load("se_obs_beta")


#Flowering function for The survival-growth kernel
flr0 <- function(x, a) {
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = unique(x),
    age0 = a)
  mu <- plogis(predict(Flowglm1, newdata = fake_data, allow.new.levels = T, type = "link"))
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}

#Flowering function for fecundity fyx0 - same function the difference is in the format of the output
flr1 <- function(x, a, Beta) {
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = unique(x),
    age0 = a)
  mu <- plogis(predict(Flowglm1, fake_data, allow.new.levels = T, type = "link") + Beta)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(matrix(mu2$V1, MatrixDim, MatrixDim, byrow = T))
}

# Growthglm1 <- fitme(log(Size1Mars) ~ 1 + poly(log(Size0Mars),4), data=growthdata, resid.model = ~ 1)
#  
# Growthglm1 <- fitme(sqrt(Size1Mars) ~ 1 + poly(Size0Mars,4) + poly(age0,3) + (Size0Mars|year) + (Size0Mars|Pop),
#                     data=growthdata,
#                     resid.model =~log(Size0Mars))

sizesObserved=sort(unique(centauree_data$Size0Mars, na.rm=T))
SD=NULL
for (i in 1:length(sizesObserved)){
  SD[i]=sd(centauree_data$Size1Mars[which(centauree_data$Size1Mars !=0 & centauree_data$Size0Mars==(sizesObserved[i]))])}
SD=SD[which(SD!=0)]
sizesObserved=sizesObserved[which(SD!=0)]
SDlm=lm(SD ~ sizesObserved)

# growthdata <- centauree_data %>%
#   filter(!is.na(Size1Mars),Size1Mars!=0,!is.na(age0)) %>%
#   mutate(logSize1Mars = log(Size1Mars),
#          logSize0Mars = log(Size0Mars))
# sizesObserved=sort(unique(growthdata$logSize0Mars, na.rm=T))
# SD=NULL
# for (i in 1:length(sizesObserved)){
#   SD[i] <- sd(growthdata$logSize1Mars[growthdata$logSize0Mars==(sizesObserved[i])])
#   }
# SDlm=lm(SD ~ sizesObserved)


#Growth function
Gyx0 <- function (y, x, a) { # Méthode midpoint
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = unique(x),
    age0 = a)
  mu <- predict(Growthglm2, fake_data, allow.new.levels = T) #Prédit les log(taille) à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)$V1 #Calcule la moyenne des log(taille)t+1
  M <- matrix(mu2, length(x), length(y), byrow = T)
  usd <- sqrt(exp(1.0462608+0.5367287*log(unique(x))))
  return(dnorm(y, mean = mu2, sd = usd)) #Proba qu'un individu soit de taille y à t+1 sachant la moyenne et la variance
}

Gyx1 <- function(ysup,yinf,x,a){ # Croissance CDF log
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = unique(x),
    age0 = a)
  mu <- exp(predict(Growthglm12, fake_data, allow.new.levels = T)) #Prédit les taille à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  usd <- sqrt(0.27795)
  usd <- sqrt(exp(-0.8872301-0.4646620*log(unique(x))))
  
  cdf.sup <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=yinf))
  
  return(cdf.sup - cdf.inf)}


Gyx2 <- function(ysup,yinf,x,a){ # Croissance CDF sans log 
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
    year = annees,
    Pop = populations,
    Size0Mars = unique(x),
    age0 = a)
  mu <- predict(Growthglm2, fake_data, allow.new.levels = T) #Prédit les taille à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  usd <- sqrt(exp(1.0462608+0.5367287*log(unique(x))))
  
  cdf.sup <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=yinf))
  # cdf.sup <- mapply(pnorm, q=ysup, MoreArgs = list(mean = mu2$V1, sd = usd))
  # cdf.inf <- mapply(pnorm, q=yinf, MoreArgs = list(mean = mu2$V1, sd = usd))
  return(cdf.sup - cdf.inf)}

#Seedling size density
densSeedl <- function(ysup, yinf) {
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  mu <- predict(Pltglm1, newdata = fake_data)[, 1]
  shape <- 1 / Pltglm1$phi
  mu <- mean(mu)
  scale <- mu * Pltglm1$phi
  cdf.sup <- pgamma(ysup, shape = shape, scale = scale)
  cdf.inf <- pgamma(yinf, shape = shape, scale = scale)
  return(cdf.sup - cdf.inf)
}

Estbl <- function(Pop=Pop, year=annees){
  fake_data <- expand.grid(
    year = year,
    Pop = Pop)
  mu <- mean(predict(Estbl))
  return(mu)
}

#Combine fecundity and flowering probability 
fyx02 <- function(x , a, Estbl, Beta = 0) {
  #flowering probability:
  p.flow <- flr1(x, a, Beta)
  #number of capitula per flowering plant
  n.captl <- numbofcap(x, a)
  #survival probability
  if(a==1){p.surv <- sx01(x,1)}
  if(a!=1){p.surv <- sx02(x,a)}
  # p.surv <- ifelse(a==1,sx01(x,a),sx02(x,a))
  # p.surv <- matrix(p.surv, MatrixDim, MatrixDim, byrow = T)
  
  Fyx0 <- p.flow * n.captl * Estbl 
  # * p.surv
  return(Fyx0)
}

fyx0 <- function(y, x , a, Estbl, Beta = 0, intervalle) {
  #flowering probability:
  p.flow <- flr1(x, a, Beta)
  #number of capitula per flowering plant
  n.captl <- numbofcap(x, a)
  #Seedlings Size distribution
  ProporSeedlSize <- densSeedl2(y,intervalle)
  #survival probability
  sx0 <- ifelse(a==1,sx01,sx02)
  p.surv <- sx0(x,a)
  
  Fyx0 <- ifelse((p.flow * n.captl * Estbl * ProporSeedlSize) >= 0,
                 p.flow * n.captl * Estbl * ProporSeedlSize * p.surv,
                 0)
  return(Fyx0)
}
```

```{r}
n.size <- 50
age <- 2
estbl <- 0.4912
Beta <- 0

K.fnc0 <- function(age, n.size, estbl, Beta=0){
  
  minsize <-  0.5
  maxsize <-  25
  maxsize1 <- 15
  h <-  (maxsize - minsize) / n.size
  h1 <-  (maxsize1 - minsize) / n.size
  b = minsize + c(0:n.size) * h
  
  #sizes for an individual of age a
  #midpoint
  ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
  #intervals
  yI.a <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.a[i,1] <- b[i]
    yI.a[i,2] <- b[i+1]
  }
  
  b1 = minsize + c(0:n.size) * h1
  #sizes for an individual of age 1
  #midpoint
  ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
  #intervals
  yI.1 <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.1[i,1] <- b1[i]
    yI.1[i,2] <- b1[i+1]
  }
  
  if(age==1){ymid.a <- ymid.1 ; yI.a <- yI.1; h <- h1}
  
  #The growth kernel
  G <- Gyx2(ysup=yI.a[,2], yinf=yI.a[,1], x=ymid.a, a=age)
  
  #The survival-growth kernel #remove flowering plants since they died
  sx0 <- ifelse(age==1,sx01,sx02)
  if(age==1){surv <- sx01(ymid.a,1)}
  if(age!=1){surv <- sx02(ymid.a,age)}
  S_flr <- surv*(1-flr0(ymid.a,age))
  # S_flr <- sx0(ymid.a,age)*(1-flr0(ymid.a,age,Beta))
  Ps <- sweep(G, MARGIN=2, S_flr, '*')
  P <- matrix(Ps, nrow=n.size)
  
  #Fecundity
  
  ProporSeedlSize <- densSeedl(ysup=yI.1[,2],yinf=yI.1[,1])
  feco <- fyx02(x = ymid.a, a = age, Estbl = estbl, Beta = Beta)
  Fec <- sweep(feco, MARGIN=1, ProporSeedlSize, '*')
  
  IPMk <- P + Fec
  return(list(M=IPMk,Pk=P,Fk=Fec))
}
```

the super matrix and growth rate
```{r}
super_matrix = function(Beta_test){
  Pk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Fk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Mk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Kernals <- array(data = 0, dim = c(AgeMax, 3, MatrixDim, MatrixDim))
  GlobMatrIPM <- array(0, dim = c((AgeMax * MatrixDim), (AgeMax * MatrixDim)))
  
  for (a in (1:(AgeMax - 1))){
    Kls <- K.fnc0(a, MatrixDim, estbl, Beta_test)
    Kernals[a, 1, , ] <- Kls$M
    Kernals[a, 2, , ] <- Kls$Pk
    Kernals[a, 3, , ] <- Kls$Fk
    # print(lambda(Kernals[a, 1, , ]))
    debut <- (a - 1) * MatrixDim + 1
    fin <- debut + (MatrixDim - 1)
    GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
    GlobMatrIPM[((fin + 1):(fin + MatrixDim)), (debut:fin)] <- Kls$Pk
  }
  
  Kls <- K.fnc0(AgeMax, MatrixDim, estbl, Beta_test)
  Kernals[AgeMax, 1, , ] <- Kls$M
  Kernals[AgeMax, 2, , ] <- Kls$Pk
  Kernals[AgeMax, 3, , ] <- Kls$Fk
  # print(lambda(Kernals[a,1,,]))
  debut <- (AgeMax - 1) * MatrixDim + 1
  fin <- debut + (MatrixDim - 1)
  GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
  GlobMatrIPM[(debut:fin), (debut:fin)] <- Kls$Pk
  
  return(GlobMatrIPM)
}
```

```{r}
MatrixIPM <- super_matrix(0) # growthrate values reported in table 1
print(lambda(MatrixIPM))
A <- eigen.analysis(MatrixIPM)
N <- matrix(ncol = 8, nrow = 50)
for (i in 1:8){
  N[,i] <- A$stable.stage[(50*(i-1)+1):(50*i)]
}
# N <- 100000*N
```

```{r}
minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b = minsize + c(0:n.size) * h

#sizes for an individual of age a
#midpoint
ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
#intervals
yI.a <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.a[i,1] <- b[i]
  yI.a[i,2] <- b[i+1]
}

b1 = minsize + c(0:n.size) * h1
#sizes for an individual of age 1
#midpoint
ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
#intervals
yI.1 <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.1[i,1] <- b1[i]
  yI.1[i,2] <- b1[i+1]
}

V <- function(age){
  return(N[(50*(age-1)+1):(50*age)])
}

Mf <- function(i,a,j){
  return(MatrixIPM[])
}
```


```{r,}
N1t1 <- sum(MatrixIPM[i,1:50] * V(1))
N3t0 <- 0
for(a in 2:8){
  N3t0 <- N3t0 + sum(V(a)*flr0(ymid.a,a))
  for (i in 1:MatrixDim){
    N1t1 <- N1t1 + sum(MatrixIPM[i,((a-1)*50+1):(a*50)] * V(a))
  }
}
fs0 <- N1t1/N3t0
fs0
```

```{r}
N2t1 <- sum(MatrixIPM[i,1:50] * V(1))
N1t0 <- sum(V(1))
for (i in 1:MatrixDim){
  N2t1 <- N2t1 + sum(MatrixIPM[50+i,] * V(1) * (1-flr0(ymid.a,2)))
}
s1_1a1 <- N2t1/N1t0
s1_1a1
```

```{r}
N2t1 <- sum(MatrixIPM[i,1:50] * V(1))
N1t0 <- sum(V(1))
for (i in 1:MatrixDim){
  N2t1 <- N2t1 + sum(MatrixIPM[50+i,] * V(1) * flr0(ymid.a,2))
}
s1_a1 <- N2t1/N1t0
s1_a1
```

```{r}
N2t1 <- 0
N2t0 <- sum(V(8)*(1-flr0(ymid.a,8)))
for(a in 2:7){
  N2t0 <- N2t0 + sum(V(a)*(1-flr0(ymid.a,a)))
  for (i in 1:MatrixDim){
    N2t1 <- N2t1 + sum(MatrixIPM[(a*50+i),] * V(a) * (1-flr0(ymid.a,a+1)))
  }
}
s2_1a2 <- N2t1/N2t0
s2_1a2
```

```{r}
N2t1 <- 0
N2t0 <- sum(V(8)*(1-flr0(ymid.a,8)))
for(a in 2:7){
  N2t0 <- N2t0 + sum(V(a)*(1-flr0(ymid.a,a)))
  for (i in 1:MatrixDim){
    N2t1 <- N2t1 + sum(MatrixIPM[(a*50+i),] * V(a) * flr0(ymid.a,a+1))
  }
}
s2_a2 <- N2t1/N2t0
s2_a2
```

```{r}
IPM <- matrix(c(0,s1_1a1,s1_a1,0,s2_1a2,s2_a2,fs0,0,0),nrow=3,ncol=3)

load("matrice_globale.RData")
MPM <- A
rm(A)

IPM
MPM
```

# Par années


## Models

Load the false data set which contains the predicted responses for survival probability, flowering probability, number of capitules (fecondity) and growth. There is a value of each of these response for each combination of size, age, population and year.
```{r}
## Creation of fake data
AgeMax <- 8
year <- 1995:2022
Pop <- c("Po","Au","Pe","E1","E2","Cr")
taille_range <- seq(0.5, 25, by = 0.5) 
age_range <- 1:AgeMax
# 
# fake_data <- expand.grid(
#   year = annees,
#   Pop = populations,
#   Size0Mars = NA,
#   age0 = NA)
# 

# Pop=gl(6,length(year),labels = Pop)
# rer$Index=1:(length(year)*6)
rer=data.frame(year)
fake_data=do.call("rbind", replicate(MatrixDim, rer, simplify = FALSE))
# fake_data=fake_data[order(fake_data$Index),]
fake_data$Size0Mars=NA
fake_data$age0=NA
fake_data$Pop <- NA

rm(rer, Pop, year)

load("ModelsAIC")

growthdata <- centauree_data[!is.na(centauree_data$Size1Mars), ]
growthdata <- growthdata[growthdata$Size1Mars != 0, ]
growthdata <- growthdata[!is.na(growthdata$age0),]
```

Creation of the functions for the kernels. 
```{r}
#Fecundity
numbofcap <- function(x, a, year) {
  fake_data <- expand.grid(
  year = year,
  Size0Mars = unique(x),
  age0 = a)
  mu <- exp(predict(Cptlglm1, fake_data))
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  nbcap <- ifelse(mu2$V1>=0,mu2$V1,0)
  return(matrix(nbcap, length(x), length(x), byrow = T))
}


#Survival Probability for age 1
sx01 <- function(x,a,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- predict(Survglm11, fake_data, allow.new.levels = T)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}

#Survival Probability for age > 1
sx02 <- function(x,a,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- predict(Survglm12, fake_data, allow.new.levels = T)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}


#Flowering Probability
#Beta = will be used to modify the intercept of the flowering function
#obs_beta = observed value
#extract Beta0
load("obs_beta")
load("se_obs_beta")


#Flowering function for The survival-growth kernel
flr0 <- function(x, a, Beta,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- plogis(predict(Flowglm1, newdata = fake_data, allow.new.levels = T, type = "link") + Beta)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(mu2$V1)
}

#Flowering function for fecundity fyx0 - same function the difference is in the format of the output
flr1 <- function(x, a, Beta,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- plogis(predict(Flowglm1, fake_data, allow.new.levels = T, type = "link") + Beta)
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean)
  return(matrix(mu2$V1, MatrixDim, MatrixDim, byrow = T))
}

sizesObserved=sort(unique(centauree_data$Size0Mars, na.rm=T))
SD=NULL
for (i in 1:length(sizesObserved)){
  SD[i]=sd(centauree_data$Size1Mars[which(centauree_data$Size1Mars !=0 & centauree_data$Size0Mars==(sizesObserved[i]))])}
SD=SD[which(SD!=0)]
sizesObserved=sizesObserved[which(SD!=0)]
SDlm=lm(SD ~ sizesObserved)

#Growth function
Gyx0 <- function (y, x, a,Pop,year) { #Méthode midpoint
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- predict(Growthglm1, fake_data, allow.new.levels = T, residVar=TRUE) #Prédit les log(taille) à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean) #Calcule la moyenne des log(taille)t+1
  M <- matrix(mu2$V1, MatrixDim, MatrixDim, byrow = T)
  Sdev <- mean(residVar(Growthglm1))
  return(dnorm(y, mean = exp(M), sd = exp(Sdev))) #Proba qu'un individu soit de taille y à t+1 sachant la moyenne et la variance
}
# summary(Gyx0(y,x,a))

annees <- 1995:2022
populations <- c("Po","Au","Pe","E1","E2","Cr")

Gyx1 <- function(ysup,yinf,x,a,Pop,year){ # Croissance CDF log
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- exp(predict(Growthglm12, fake_data, allow.new.levels = T)) #Prédit les taille à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  usd <- sqrt(0.27795)
  # usd <- sqrt(exp(-0.8872301-0.4646620*log(unique(x))))

  cdf.sup <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
}

Gyx2 <- function(ysup,yinf,x,a,Pop,year){ # Croissance CDF sans log
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  age0 = a)
  mu <- predict(Growthglm2, fake_data, allow.new.levels = T) #Prédit les taille à t+1 possibles 
  mu2 <- aggregate(mu, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  usd <- sqrt(exp(1.0462608+0.5367287*log(unique(x))))

  cdf.sup <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = mu2$V1, sd = usd, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
}

densSeedl <- function(ysup, yinf,Pop,year) {
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  fake_data <- expand.grid(
  year = year,
  Pop = Pop)
  mu <- predict(Pltglm1, newdata = fake_data)[, 1]
  shape <- 1 / Pltglm1$phi
  mu <- mean(mu)
  scale <- mu * Pltglm1$phi
  cdf.sup <- pgamma(ysup, shape = shape, scale = scale)
  cdf.inf <- pgamma(yinf, shape = shape, scale = scale)
  return(cdf.sup - cdf.inf)
}

Pop <- populations
year <- annees
Estbli <- function(Pop, year){
  fake_data <- expand.grid(
  year = year,
  Pop = Pop)
  mu <- mean(predict(Estbglm1, newdata= fake_data))
  return(mu)
}
#Combine fecundity and flowering probability 
fyx0 <- function(x , a, Beta = 0, Pop,year) {
  #flowering probability:
  p.flow <- flr1(x, a, Beta,Pop,year)
  #number of capitula per flowering plant
  n.captl <- numbofcap(x, a, year)
  #survival probability
  sx0 <- ifelse(a==1,sx01,sx02)
  p.surv <- sx0(x,a,Pop,year)
  
  #Establishment rate
  Estbl <- Estbli(Pop, year)
  
  feco <- p.flow * n.captl * Estbl 
  #* p.surv
  return(feco)
}
```

```{r}
n.size <- 50

K.fnc0 <- function(age, n.size, Beta=0, Pop,year){
  
  minsize <-  0.5
  maxsize <-  25
  maxsize1 <- 15
  h <-  (maxsize - minsize) / n.size
  h1 <-  (maxsize1 - minsize) / n.size
  b = minsize + c(0:n.size) * h

  #sizes for an individual of age a
  #midpoint
  ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
  #intervals
  yI.a <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.a[i,1] <- b[i]
    yI.a[i,2] <- b[i+1]
  }
  
  b1 = minsize + c(0:n.size) * h1
  #sizes for an individual of age 1
  #midpoint
  ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
  #intervals
  yI.1 <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.1[i,1] <- b1[i]
    yI.1[i,2] <- b1[i+1]
  }
  
  if(age==1){ymid.a <- ymid.1 ; yI.a <- yI.1; h <- h1}
  
  #The growth kernel
  G <- Gyx2(ysup=yI.a[,2], yinf=yI.a[,1], x=ymid.a, a=age, Pop=Pop, year=year)

    #The survival-growth kernel #remove flowering plants since they died
  sx0 <- ifelse(age==1,sx01,sx02)
  S_flr <- sx0(ymid.a,age, Pop=Pop, year=year)*(1-flr0(ymid.a,age,Beta, Pop=Pop, year=year))
  Ps <- sweep(G, MARGIN=2, S_flr, '*')
  P <- matrix(Ps, nrow=n.size)
  
  #Fecundity
  
  ProporSeedlSize <- densSeedl(ysup=yI.1[,2],yinf=yI.1[,1], Pop=Pop, year=year)
  feco <- fyx0(x = ymid.a, a = age, Beta = Beta, Pop=Pop, year=year)
  Fec <- sweep(feco, MARGIN=1, ProporSeedlSize, '*')

  IPMk <- P + Fec
  return(list(M=IPMk,Pk=P,Fk=Fec))
}
```

  
  the super matrix and growth rate
```{r}
super_matrix = function(Beta_test,Pop,year){
  Pk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Fk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Mk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Kernals <- array(data = 0, dim = c(AgeMax, 3, MatrixDim, MatrixDim))
  GlobMatrIPM <- array(0, dim = c((AgeMax * MatrixDim), (AgeMax * MatrixDim)))
  
  for (a in (1:(AgeMax - 1))){
    Kls <- K.fnc0(a, MatrixDim, Beta_test, Pop, year)
    Kernals[a, 1, , ] <- Kls$M
    Kernals[a, 2, , ] <- Kls$Pk
    Kernals[a, 3, , ] <- Kls$Fk
    # print(lambda(Kernals[a, 1, , ]))
    debut <- (a - 1) * MatrixDim + 1
    fin <- debut + (MatrixDim - 1)
    GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
    GlobMatrIPM[((fin + 1):(fin + MatrixDim)), (debut:fin)] <- Kls$Pk
  }
  
  Kls <- K.fnc0(AgeMax, MatrixDim, Beta_test, Pop, year)
  Kernals[AgeMax, 1, , ] <- Kls$M
  Kernals[AgeMax, 2, , ] <- Kls$Pk
  Kernals[AgeMax, 3, , ] <- Kls$Fk
  # print(lambda(Kernals[a,1,,]))
  debut <- (AgeMax - 1) * MatrixDim + 1
  fin <- debut + (MatrixDim - 1)
  GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
  GlobMatrIPM[(debut:fin), (debut:fin)] <- Kls$Pk
  
  return(GlobMatrIPM)
}
```


Par années et pop
```{r}
Pop <- c("Po","Au","Pe","E1","E2","Cr")
annees <- 1995:2022
lam <- expand.grid(year=annees,Pop=Pop,lambda=NA)

for(i in 1:6) {
  for (j in 1:length(annees)) {
    obs_lambda <- super_matrix(0, Pop[i], annees[j]) # growthrate values reported in table 1
    # print(obs_lambda)
    lam$lambda[lam$year==annees[j] & lam$Pop==Pop[i]] <- obs_lambda
  }
}

mean(lam$lambda)
print(lam)
ggplot(data=lam,aes(x=year,y=lambda))+
  geom_abline(intercept=1,slope=0)+
  geom_point(aes(colour = as.factor(Pop),shape = as.factor(Pop)))+
  geom_line(aes(colour = as.factor(Pop)))+
  theme_bw()+
  ylim(0.2,1.8)

print(obs_beta)

mean_size <- mean(centauree_data_complet$Size0Mars[which(centauree_data_complet$Flowering0 == 1)])
sd_size <- sd(centauree_data_complet$Size0Mars[which(centauree_data_complet$Flowering0 == 1)])
print(mean_size)
print(sd_size)

mean_age <- mean(centauree_data_complet$age0[which(centauree_data_complet$Flowering0 == 1)], na.rm = T)
sd_age <- sd(centauree_data_complet$age0[which(centauree_data_complet$Flowering0 == 1)], na.rm = T)

print(mean_age)
print(sd_age)
```

Par années, Pop pooled
```{r}
Pop <- c("Po","Au","Pe","E1","E2","Cr")
annees <- 1995:2022
lam <- expand.grid(year=annees,lambda=NA)


for (j in 1:length(annees)) {
  obs_lambda <- lambda(super_matrix(0, Pop, annees[j])) # growthrate values reported in table 1
  print(obs_lambda)
  lam$lambda[lam$year==annees[j]] <- obs_lambda
  }

print(lam)

lamMPM <- GrowthRateMPM %>% 
  filter(Annee != 1994) %>% 
  group_by(Annee) %>% 
  summarize(year = Annee,
            lambdaMPM=mean(E1,E2,Cr,Au,Pe,Po)) %>% 
  select(!Annee)

lam %>% 
  left_join(lamMPM, by = join_by(year==year)) %>% 
  ggplot(aes(x=year,lambda))+
    geom_abline(intercept=1,slope=0,linetype=2)+
    geom_point(color="red")+
    geom_line(color="red")+
    geom_point(aes(y=lambdaMPM))+
    geom_line(aes(y=lambdaMPM))+
    theme_bw()+
    labs(title = "MPM en noir ; IPM en rouge")+
    ylim(0.2,2.5)

  geom_smooth(color="pink",se=FALSE)
```


```{r}
MatrixIPM <- list()
Popu <- Pop
for(i in 1:6) {
  for (k in 1:length(annees)) {
    obs_lambda <- super_matrix(0, Popu[i], annees[k]) # growthrate values reported in table 1
    MatrixIPM[[k]] <- obs_lambda
  }
}
```

```{r}
minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b = minsize + c(0:n.size) * h

#sizes for an individual of age a
#midpoint
ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
#intervals
yI.a <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.a[i,1] <- b[i]
  yI.a[i,2] <- b[i+1]
}

b1 = minsize + c(0:n.size) * h1
#sizes for an individual of age 1
#midpoint
ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
#intervals
yI.1 <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.1[i,1] <- b1[i]
  yI.1[i,2] <- b1[i+1]
}
```


```{r}
IPMyear <- list()
for (k in 1:length(annees)) {
  year <- annees[k]
  M <- MatrixIPM[[k]]
  A <- eigen.analysis(M)
  N <- matrix(ncol = 8, nrow = 50)
  for (i in 1:8) {
    N[, i] <- A$stable.stage[(50 * (i - 1) + 1):(50 * i)]
  }
  V <- function(age) {
    return(N[(50 * (age - 1) + 1):(50 * age)])
  }
  
  N1t1 <- sum(M[i, 1:50] * V(1))
  N3t0 <- 0
  for (a in 2:8) {
    N3t0 <- N3t0 + sum(V(a) * flr0(ymid.a, a, 0,Pop, year))
    for (i in 1:MatrixDim) {
      N1t1 <- N1t1 + sum(M[i, ((a - 1) * 50 + 1):(a * 50)] * V(a))
    }
  }
  fs0 <- N1t1 / N3t0
  fs0
  
  N2t1 <- sum(M[i, 1:50] * V(1))
  N1t0 <- sum(V(1))
  for (i in 1:MatrixDim) {
    N2t1 <- N2t1 + sum(M[50 + i, ] * V(1) * (1 - flr0(ymid.a, 2,0, Pop, year)))
  }
  s1_1a1 <- N2t1 / N1t0
  s1_1a1
  
  N2t1 <- sum(M[i, 1:50] * V(1))
  N1t0 <- sum(V(1))
  for (i in 1:MatrixDim) {
    N2t1 <- N2t1 + sum(M[50 + i, ] * V(1) * flr0(ymid.a, 2, 0,Pop, year))
  }
  s1_a1 <- N2t1 / N1t0
  s1_a1
  
  N2t1 <- 0
  N2t0 <- sum(V(8) * (1 - flr0(ymid.a, 8,0, Pop, year)))
  for (a in 2:7) {
    N2t0 <- N2t0 + sum(V(a) * (1 - flr0(ymid.a, a, 0,Pop, year)))
    for (i in 1:MatrixDim) {
      N2t1 <- N2t1 + sum(M[(a * 50 + i), ] * V(a) * (1 - flr0(ymid.a, a+1,0, Pop, year)))
    }
  }
  s2_1a2 <- N2t1 / N2t0
  s2_1a2
  
  N2t1 <- 0
  N2t0 <- sum(V(8) * (1 - flr0(ymid.a, 8, 0,Pop, year)))
  for (a in 2:7) {
    N2t0 <- N2t0 + sum(V(a) * (1 - flr0(ymid.a, a,0, Pop, year)))
    for (i in 1:MatrixDim) {
      N2t1 <- N2t1 + sum(M[(a * 50 + i), ] * V(a) * flr0(ymid.a, a + 1,0, Pop, year))
    }
  }
  s2_a2 <- N2t1 / N2t0
  s2_a2
  
  IPMyear[[k]] <- matrix(c(0, s1_1a1, s1_a1, 0, s2_1a2, s2_a2, fs0, 0, 0),
                     nrow = 3,
                     ncol = 3)
}
save(IPMyear,file = "IPMEstimatedMatrix.RData")

```


```{r}
load("IPMEstimatedMatrix.RData")
load("matricesperyear.RData")
MPMyear <- matrices
rm(matrices)

IPMyear <- IPM
# MPMyear
```

```{r}
i <- NULL
m <- NULL
for(k in 1:(length(annees)-1)){
# image2(IPMyear[[k]]);image2(MPMyear[[k]])
  i[k] <- IPMyear[[k]][1,3] 
  m[k] <- MPMyear[[k]][1,3]
}
plot(x=annees[-length(annees)],i,ylim=c(0,30),type="l");lines(x=annees[-length(annees)],m,col="red")
```

