---
  title: "Optimal size at flowering"
author: "Loïc Pages"
date: "`r Sys.Date()`"
output:
  pdf_document: default
word_document: default
html_document:
  df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/media/loic/Commun/0Travail/2025 IPM Centauree ISEM/2025 Stage ISEM/Code")
library(splines)
library(tidyverse)
library(spaMM)
library(popbio)
```


# Initialisation

```{r}
IPM_data <- read.csv("Data/newdata.csv")

MatrixDim <- 50
RecrRate <- 0.338 #recruitment rate

#Forcer l'age maximal à 8
AgeMax <- 8
IPM_data$Age[IPM_data$Age > AgeMax] <- AgeMax

centauree_data <- IPM_data[!is.na(IPM_data$Size0Mars) & !is.na(IPM_data$Age),]
```

```{r}
load("IPM/Kernals implementation/IPMKernalPop_AIC.RData")
load("IPM/Kernals implementation/IPMKernalYear_AIC.RData")
load("IPM/Kernals implementation/IPMKernalPopYear_AIC.RData")
```

Load the selected models

```{r}
load("IPM/ModelsAIC.RData")
```

Creation of the functions for the kernels.

```{r}
annees <- 1995:2022
populations <- c("Po","Au","Pe","E1","E2","Cr")

#Fecundity
numbofcap <- function(x, a,year) {
  fake_data <- expand.grid(
  year = year,
  Size0Mars = unique(x),
  Age = a)
  predi <- exp(predict(Cptlglm1, fake_data)) # prédiction du nombre de capitules produits pour iun individu de taille x, d'age a, à l'année year
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) #transforme le vecteur de prédiction en un vecteur au bon format, tout en faisant une moyenne si besoin (dans le cas où on pool les populations)
  
  meanpredi$V1[meanpredi$Group.1<5] <- 0
  nbcap <- ifelse(meanpredi$V1>=0,meanpredi$V1,0)
  return(matrix(nbcap, length(x), length(x), byrow = T))
}


#Survival Probability for age 1
sx01 <- function(x,a,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- suppressWarnings(predict(Survglm11, fake_data, allow.new.levels = T))
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(meanpredi$V1)
}

#Survival Probability for age > 1
sx02 <- function(x,a,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- suppressWarnings(predict(Survglm12, fake_data, allow.new.levels = T))
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(meanpredi$V1)
}


#Flowering Probability
#Beta = will be used to modify the intercept of the flowering function
#obs_beta = observed value
#extract Beta0
load("obs_beta")
load("se_obs_beta")


#Flowering function for The survival-growth kernel
flr0 <- function(x, a, Beta=0,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- plogis(predict(Flowglm1, newdata = fake_data, allow.new.levels = T, type = "link") + Beta)
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(meanpredi$V1)
}

#Flowering function for fecundity fyx0 - same function the difference is in the format of the output
flr1 <- function(x, a, Beta,Pop,year) {
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- plogis(predict(Flowglm1, fake_data, allow.new.levels = T, type = "link") + Beta)
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean)
  return(matrix(meanpredi$V1, MatrixDim, MatrixDim, byrow = T))
}


#Growth (cumulative ditribution function)

## Gaussian
Gyx1 <- function(ysup,yinf,x,a,Pop,year){
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- predict(Growthglm1, fake_data, allow.new.levels = T) # Prédit les taille à t+1 possibles
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) # Transforme le vecteur de prédiction en un vecteur au bon format, tout en faisant une moyenne si besoin (dans le cas où on pool les populations)
  varpredi <- sqrt(exp(1.1013346 + 0.5553254*log(unique(x)) - 0.1231723*log(a))) # Variance résiduelle du modèle

  cdf.sup <- mapply(pnorm, mean = meanpredi$V1, sd = varpredi, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = meanpredi$V1, sd = varpredi, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
}

## Gamma(log)
Gyx2 <- function(ysup,yinf,x,a,Pop,year){
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- predict(Growthglm1gamma, fake_data, allow.new.levels = T) #Prédit les taille à t+1 possibles 
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) 
  phi <- exp(-1.07986 - 0.57167*log(unique(x)) + 0.06724*log(a))
  varpredi <- phi*meanpredi$V1

  cdf.sup <- mapply(pnorm, mean = meanpredi$V1, sd = varpredi, MoreArgs = list(q=ysup))
  cdf.inf <- mapply(pnorm, mean = meanpredi$V1, sd = varpredi, MoreArgs = list(q=yinf))
  return(cdf.sup - cdf.inf)
}

## Gaussian improved (cox-box power + random effect on residual variance)
Gyx3 <- function(ysup,yinf,x,a,Pop,year){  
  sizepow <- 0.4343354
  yinfpow <- yinf**sizepow
  ysuppow <- ysup**sizepow
  yinfpow[1] <- -Inf
  ysuppow[length(ysuppow)] <- Inf
  
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Size0Mars = unique(x),
  Age = a)
  predi <- suppressWarnings(predict(Growthglm2, fake_data, allow.new.levels = T)) #Prédit les taille^pow à t+1 possibles 
  meanpredi <- aggregate(predi, list(fake_data$Size0Mars), mean) #Calcule la moyenne des taillet+1
  
  resvar <- predict(Growthglm2$resid_fit, newdata = fake_data, allow.new.levels = T)
  meanresvar <- aggregate(resvar[,1], list(fake_data$Size0Mars), mean)$x
  
  # cdf.sup <- mapply(pnorm, mean = meanpredi$V1, sd = sqrt(meanvarpredi), MoreArgs = list(q=ysup))
  # cdf.inf <- mapply(pnorm, mean = meanpredi$V1, sd = sqrt(meanvarpredi), MoreArgs = list(q=yinf))
  cdf.sup <- mapply(pnorm, mean = meanpredi$V1, sd = sqrt(meanresvar), MoreArgs = list(q=ysuppow))
  cdf.inf <- mapply(pnorm, mean = meanpredi$V1, sd = sqrt(meanresvar), MoreArgs = list(q=yinfpow))
  return(cdf.sup - cdf.inf)
}


densSeedl <- function(ysup, yinf,Pop,year) {
  yinf[1] <- -Inf
  ysup[length(ysup)] <- Inf
  fake_data <- expand.grid(
  year = year,
  Pop = Pop)
  
  mu <- predict(Pltglm1, newdata = fake_data)[, 1]
  shape <- 1 / Pltglm1$phi
  mu <- mean(mu)
  scale <- mu * Pltglm1$phi
  cdf.sup <- pgamma(ysup, shape = shape, scale = scale)
  cdf.inf <- pgamma(yinf, shape = shape, scale = scale)
  return(cdf.sup - cdf.inf)
}

Pop <- populations
year <- annees

Estbli <- function(Pop, year){
  fake_data <- expand.grid(
  year = year,
  Pop = Pop,
  Capitule = 1:200)
  plt <- predict(Estbglm2, newdata = fake_data)[,1]
  mu <- mean(plt/fake_data$Capitule)
  return(mu)
}


#Combine fecundity and flowering probability 
fyx0 <- function(x , a, Beta = 0, Pop,year) {
  #flowering probability:
  p.flow <- flr1(x, a, Beta,Pop,year)
  #number of capitula per flowering plant
  n.captl <- numbofcap(x, a,year)
  #survival probability
  sx0 <- ifelse(a==1,sx01,sx02)
  p.surv <- sx0(x,a,Pop,year)
  
  #Establishment rate
  Estbl <- Estbli(Pop, year)
  # Estbl <- RecrRate
  
  feco <- p.flow * n.captl * Estbl 
  #* p.surv
  return(feco)
}
```

```{r}
K.fnc0 <- function(age, n.size, Beta=0, Pop,year){
  
  minsize <-  0.5
  maxsize <-  25
  maxsize1 <- 15
  h <-  (maxsize - minsize) / n.size
  h1 <-  (maxsize1 - minsize) / n.size
  b = minsize + c(0:n.size) * h

  #sizes for an individual of age a
  #midpoint
  ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
  #intervals
  yI.a <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.a[i,1] <- b[i]
    yI.a[i,2] <- b[i+1]
  }
  
  b1 = minsize + c(0:n.size) * h1
  #sizes for an individual of age 1
  #midpoint
  ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
  #intervals
  yI.1 <- matrix(nrow = n.size, ncol = 2)
  for (i in 1:n.size){
    yI.1[i,1] <- b1[i]
    yI.1[i,2] <- b1[i+1]
  }
  
  if(age==1){ymid.a <- ymid.1
  # yI.a <- yI.1 
  h <- h1}
  
  #The growth kernel
  # G <- Gyx1(ysup=yI.a[,2], yinf=yI.a[,1], x=ymid.a, a=age, Pop=Pop, year=year)
  # G <- Gyx2(ysup=yI.a[,2], yinf=yI.a[,1], x=ymid.a, a=age, Pop=Pop, year=year)
  G <- Gyx3(ysup=yI.a[,2], yinf=yI.a[,1], x=ymid.a, a=age, Pop=Pop, year=year)

  #The survival-growth kernel #remove flowering plants since they died
  sx0 <- ifelse(age==1,sx01,sx02)
  S_flr <- sx0(ymid.a,age, Pop=Pop, year=year)*(1-flr0(ymid.a,age,Beta, Pop=Pop, year=year))
  Ps <- sweep(G, MARGIN=2, S_flr, '*')
  P <- matrix(Ps, nrow=n.size)
  
  #Fecundity
  
  ProporSeedlSize <- densSeedl(ysup=yI.1[,2],yinf=yI.1[,1], Pop=Pop, year=year)
  feco <- fyx0(x = ymid.a, a = age, Beta = Beta, Pop=Pop, year=year)
  Fec <- sweep(feco, MARGIN=1, ProporSeedlSize, '*')

  IPMk <- P + Fec
  return(list(M=IPMk,Pk=P,Fk=Fec))
}

super_matrix = function(Beta_test,Pop,year){
  Pk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Fk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Mk <- matrix(data = NA, nrow = MatrixDim, ncol = MatrixDim)
  Kernals <- array(data = 0, dim = c(AgeMax, 3, MatrixDim, MatrixDim))
  GlobMatrIPM <- array(0, dim = c((AgeMax * MatrixDim), (AgeMax * MatrixDim)))
  
  for (a in (1:(AgeMax - 1))){
    Kls <- K.fnc0(a, MatrixDim, Beta_test, Pop, year)
    Kernals[a, 1, , ] <- Kls$M
    Kernals[a, 2, , ] <- Kls$Pk
    Kernals[a, 3, , ] <- Kls$Fk
    # print(lambda(Kernals[a, 1, , ]))
    debut <- (a - 1) * MatrixDim + 1
    fin <- debut + (MatrixDim - 1)
    GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
    GlobMatrIPM[((fin + 1):(fin + MatrixDim)), (debut:fin)] <- Kls$Pk
  }
  
  Kls <- K.fnc0(AgeMax, MatrixDim, Beta_test, Pop, year)
  Kernals[AgeMax, 1, , ] <- Kls$M
  Kernals[AgeMax, 2, , ] <- Kls$Pk
  Kernals[AgeMax, 3, , ] <- Kls$Fk
  # print(lambda(Kernals[a,1,,]))
  debut <- (AgeMax - 1) * MatrixDim + 1
  fin <- debut + (MatrixDim - 1)
  GlobMatrIPM[1:MatrixDim, (debut:fin)] <- Kls$Fk
  GlobMatrIPM[(debut:fin), (debut:fin)] <- Kls$Pk
  
  return(GlobMatrIPM)
}
```

```{r}
minsize <-  0.5
maxsize <-  25
maxsize1 <- 15
n.size <- MatrixDim
h <-  (maxsize - minsize) / n.size
h1 <-  (maxsize1 - minsize) / n.size
b = minsize + c(0:n.size) * h

#sizes for an individual of age a
#midpoint
ymid.a = 0.5 * (b[1:n.size] + b[2:(n.size + 1)])
#intervals
yI.a <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.a[i,1] <- b[i]
  yI.a[i,2] <- b[i+1]
}

b1 = minsize + c(0:n.size) * h1
#sizes for an individual of age 1
#midpoint
ymid.1 = 0.5 * (b1[1:n.size] + b1[2:(n.size + 1)])
#intervals
yI.1 <- matrix(nrow = n.size, ncol = 2)
for (i in 1:n.size){
  yI.1[i,1] <- b1[i]
  yI.1[i,2] <- b1[i+1]
} 
```

# Optimal size at flowering per population

Calcul et plot du taux de croissance (et de la taille optimale correspondante) en fonction de la variation du beta
```{r,warning=FALSE}
load("IPM/obs_beta.RData")
load("IPM/se_obs_beta.RData")
minbeta <- -15
maxbeta <- 15
beta <- c(seq(-15,-10,5),seq(-8,-4,2),seq(-3,5,0.5),seq(6,15,2))
lam_pop <- matrix(nrow=6,ncol=length(beta))
mu_sizes_pop <- matrix(nrow=6,ncol=length(beta))
lam_obs <- NA

for (p in 1:6){
  print(p)
  A <- eigen.analysis(IPMKernalPop_AIC[[p]])
  N <- matrix(ncol = 8, nrow = 50)
  for (i in 1:8) {
    N[, i] <- A$stable.stage[(50 * (i - 1) + 1):(50 * i)]
  }
  W <- function(age) {
    return(N[(50 * (age - 1) + 1):(50 * age)])
  }
  pop <- populations[p]

  for (i in 1:length(beta)){
    lam_pop[p,i] <- lambda(super_matrix(beta[i],pop,year=annees))
    if(beta[i]==0){lam_obs[p] <- lam_pop[p,i]}
    
    mu_size <- sum(ymid.1*W(1)*flr0(ymid.1,1,beta[i],pop,year=annees))
    # mu_a <- sum(W(1)*flr0(ymid.1,1,pop,year=annees))
    mu2 <- sum(W(1)*flr0(ymid.1,1,beta[i],pop,year=annees))
    for(a in 2:8){
      mu_size <- mu_size + sum(ymid.a*W(a)*flr0(ymid.a,a,beta[i],pop,year=annees))
      # mu_a <- mu_a + sum(a*W(a)*flr0(ymid.a,a,pop,year=annees))
      mu2 <- mu2 + sum(W(a)*flr0(ymid.a,a,beta[i],pop,year=annees))
    }
    mu_sizes_pop[p,i] <- mu_size/mu2
    # mu_a <- mu_a/mu2
  }
}
```


```{r}
for (p in 1:6){
pos <- order(lam_pop[p,1:length(beta)])
maxibeta <- beta[pos[length(pos)]]

plot(x=beta+obs_beta,y=lam_pop[p,],type = 'line') ;abline(v=obs_beta-se_obs_beta,lty=2) ; abline(v=obs_beta+se_obs_beta,lty=2) ;abline(v=maxibeta+obs_beta,col="red") ; abline(v=obs_beta,col="green4",lty=2) ; abline(h=lam_obs[p],lty=2,col="green4"); abline(h=max(lam_pop[p,1:length(beta)]),col="red",lty=3) ; axis(side = 3,labels = round(mu_sizes_pop[p,], 2), at=beta+obs_beta) ; title(populations[p])
}
```


# Optimal size at flowering per year
```{r}
functionlambda <- function(beta){
return(lambda(super_matrix(beta, Pop=populations,year=1995)))
  }

optimize(f = functionlambda,maximum = TRUE,lower=-25,upper=5,tol=0.001)
```

Calcul et plot du taux de croissance (et de la taille optimale correspondante) en fonction de la variation du beta
```{r,warning=FALSE}
load("IPM/obs_beta.RData")
load("IPM/se_obs_beta.RData")
minbeta <- -15
maxbeta <- 15
beta <- c(seq(-15,-10,5),seq(-8,-4,2),seq(-3,5,0.5),seq(6,15,2))
# beta <- c(seq(minbeta,maxbeta,2))
lam_year <- matrix(nrow=length(annees),ncol=length(beta))
mu_sizes_year <- matrix(nrow=length(annees),ncol=length(beta))
lam_obs <- NA

for (yr in 1:length(annees)){
  print(yr)
  A <- eigen.analysis(IPMKernalYear_AIC[[yr]])
  N <- matrix(ncol = 8, nrow = 50)
  for (i in 1:8) {
    N[, i] <- A$stable.stage[(50 * (i - 1) + 1):(50 * i)]
  }
  W <- function(age) {
    return(N[(50 * (age - 1) + 1):(50 * age)])
  }
  year <- annees[yr]

  for (i in 1:length(beta)){
    lam_year[yr,i] <- lambda(super_matrix(beta[i],Pop=populations,year=year))
    if(beta[i]==0){lam_obs[yr] <- lam_year[yr,i]}
    
    mu_size <- sum(ymid.1*W(1)*flr0(ymid.1,1,beta[i],Pop=populations,year=year))
    # mu_a <- sum(W(1)*flr0(ymid.1,1,Pop=populations,year=year))
    mu2 <- sum(W(1)*flr0(ymid.1,1,beta[i],Pop=populations,year=year))
    for(a in 2:8){
      mu_size <- mu_size + sum(ymid.a*W(a)*flr0(ymid.a,a,beta[i],Pop=populations,year=year))
      # mu_a <- mu_a + sum(a*W(a)*flr0(ymid.a,a,Pop=populations,year=year))
      mu2 <- mu2 + sum(W(a)*flr0(ymid.a,a,beta[i],Pop=populations,year=year))
    }
    mu_sizes_year[yr,i] <- mu_size/mu2
    # mu_a <- mu_a/mu2
  }
}

mu_sizes_predi <- NA
for (yr in 1:length(annees)){
  A <- eigen.analysis(IPMKernalYear_AIC[[yr]])
  N <- matrix(ncol = 8, nrow = 50)
  for (i in 1:8) {
    N[, i] <- A$stable.stage[(50 * (i - 1) + 1):(50 * i)]
  }
  W <- function(age) {
    return(N[(50 * (age - 1) + 1):(50 * age)])
  }
  year <- annees[yr]
  mu_size <- sum(ymid.1*W(1)*flr0(ymid.1,1,0,Pop=populations,year=year))
    mu2 <- sum(W(1)*flr0(ymid.1,1,0,Pop=populations,year=year))
    for(a in 2:8){
      mu_size <- mu_size + sum(ymid.a*W(a)*flr0(ymid.a,a,0,Pop=populations,year=year))
      mu2 <- mu2 + sum(W(a)*flr0(ymid.a,a,0,Pop=populations,year=year))
    }
    mu_sizes_predi[yr] <- mu_size/mu2
}
save(mu_sizes_predi, file="IPM/Results/mu_sizes_predi.RData")
```


```{r}
opti_size_year <- NA
for (yr in 1:length(annees)){
pos <- order(lam_year[yr,])
maxibeta <- beta[pos[length(pos)]]
opti_size_year[yr] <- round(mu_sizes_year[yr,pos[length(pos)]], 2)

plot(x=beta+obs_beta,y=lam_year[yr,],type = 'line') ;abline(v=obs_beta-se_obs_beta,lty=2) ; abline(v=obs_beta+se_obs_beta,lty=2) ;abline(v=maxibeta+obs_beta,col="red") ; abline(v=obs_beta,col="green4",lty=2) ; abline(h=lam_obs[yr],lty=2,col="green4"); abline(h=max(lam_year[yr,]),col="red",lty=3) ; axis(side = 3,labels = round(mu_sizes_year[yr,], 2), at=beta+obs_beta) ; title(annees[yr])
}
# save(opti_size_year, file="IPM/Results/opti_size_year.RData")
```

# Optimal size at flowering per population and year

```{r}
mu_sizes_popyear <- matrix(NA, length(populations), length(annees))

for (p in 1:6) {
  for (yr in 1:length(annees)) {
    year <- annees[yr]
    pop <- populations[p]
    print(p)
    functionlambda <- function(beta) {
      return(lambda(super_matrix(beta, Pop = pop, year = year)))
    }
    opti <- optimize(
      f = functionlambda,
      maximum = TRUE,
      lower = -25,
      upper = 5,
      tol = 0.001)
    beta <- opti$maximum

    A <- eigen.analysis(IPMKernalYear_AIC[[yr]])
    N <- matrix(ncol = 8, nrow = 50)
    for (i in 1:8) {
      N[, i] <- A$stable.stage[(50 * (i - 1) + 1):(50 * i)]
    }
    W <- function(age) {
      return(N[(50 * (age - 1) + 1):(50 * age)])
    }
    mu_size <- sum(ymid.1*W(1)*flr0(ymid.1,1,beta,Pop=pop,year=year))
    mu2 <- sum(W(1)*flr0(ymid.1,1,beta,Pop=pop,year=year))
    for(a in 2:8){
      mu_size <- mu_size + sum(ymid.a*W(a)*flr0(ymid.a,a,beta,Pop=pop,year=year))
      mu2 <- mu2 + sum(W(a)*flr0(ymid.a,a,beta,Pop=pop,year=year))
    }
    mu_sizes_popyear[p,yr] <- mu_size/mu2
  }
}
save(mu_sizes_popyear, file="IPM/Results/mu_sizes_popyear.RData")
```

Calcul et plot du taux de croissance (et de la taille optimale correspondante) en fonction de la variation du beta
```{r,warning=FALSE}
load("IPM/obs_beta.RData")
load("IPM/se_obs_beta.RData")
minbeta <- -15
maxbeta <- 15
beta <- c(seq(-15,-10,5),seq(-8,-4,2),seq(-3,3,0.5),seq(4,15,5))
# beta <- c(seq(minbeta,maxbeta,2))
lam_popyear <- array(NA, dim = c(length(populations), length(annees),length(beta)))
mu_sizes_popyear <- array(NA, dim = c(length(populations), length(annees),length(beta)))
lam_obs <- matrix(NA,length(populations),length(annees))

for (p in 1:6){
for (yr in 1:length(annees)){
  print(yr)
  A <- eigen.analysis(IPMKernalPopYear_AIC[[28*(p-1)+yr]])
  N <- matrix(ncol = 8, nrow = 50)
  for (i in 1:8) {
    N[, i] <- A$stable.stage[(50 * (i - 1) + 1):(50 * i)]
  }
  W <- function(age) {
    return(N[(50 * (age - 1) + 1):(50 * age)])
  }
  year <- annees[yr]
  pop <- populations[p]

  for (i in 1:length(beta)){
    lam_popyear[p,yr,i] <- lambda(super_matrix(beta[i],Pop=pop,year=year))
    if(beta[i]==0){lam_obs[p,yr] <- lam_popyear[p,yr,i]}
    
    mu_size <- sum(ymid.1*W(1)*flr0(ymid.1,1,beta[i],Pop=pop,year=year))
    # mu_a <- sum(W(1)*flr0(ymid.1,1,Pop=pop,year=year))
    mu2 <- sum(W(1)*flr0(ymid.1,1,beta[i],Pop=pop,year=year))
    for(a in 2:8){
      mu_size <- mu_size + sum(ymid.a*W(a)*flr0(ymid.a,a,beta[i],Pop=pop,year=year))
      # mu_a <- mu_a + sum(a*W(a)*flr0(ymid.a,a,Pop=pop,year=year))
      mu2 <- mu2 + sum(W(a)*flr0(ymid.a,a,beta[i],Pop=pop,year=year))
    }
    mu_sizes_popyear[p,yr,i] <- mu_size/mu2
    # mu_a <- mu_a/mu2
  }
}
}
```


```{r}
opti_size_popyear <- matrix(NA,length(populations),length(annees))
for(p in 1:6){
for (yr in 1:length(annees)){
pos <- order(lam_popyear[p,yr,])
maxibeta <- beta[pos[length(pos)]]
opti_size_popyear[p,yr] <- round(mu_sizes_popyear[p,yr,pos[length(pos)]], 2)

plot(x=beta+obs_beta,y=lam_popyear[p,yr,],type = 'line') ;abline(v=obs_beta-se_obs_beta,lty=2) ; abline(v=obs_beta+se_obs_beta,lty=2) ;abline(v=maxibeta+obs_beta,col="red") ; abline(v=obs_beta,col="green4",lty=2) ; abline(h=lam_obs[p,yr],lty=2,col="green4"); abline(h=max(lam_popyear[p,yr,]),col="red",lty=3) ; axis(side = 3,labels = round(mu_sizes_popyear[p,yr,], 2), at=beta+obs_beta) ; title(c(populations[p],annees[yr]))
}
}
save(opti_size_popyear, file="IPM/Results/opti_size_popyear.RData")
```

